%%% File:      b08_Lists.tex
%%% Author:    Joaquín Ataz-López et al
%%% Begun:     July 2020
%%% Concluded: July 2020
%%% Contents:  Initially this chapter was to be part of Chapter 12
%              (elements and structures of documents).  But I saw that on
%              the one hand indexes affected the document globally, and on
%              the other hand, if I included this in Chapter 12, it would
%              become disproportionately long.  The chapter was written in
%              record time (three or four days), a sign that I was already
%              beginning to get a more intimate understanding of ConTeXt.
%
%%% Edited with: Emacs + AUCTeX - And at times vim + context-plugin
%%%


\startcomponent b08_Lists.tex

\environment introCTX_env.tex

\startchapter
  [
    reference=cap:toc,
    title={Tables of contents and \hbox{indices}},
    bookmark={Tables of contents and indices},
  ]

\TocChap

A table of contents and an index are a global aspect of a document.  Almost
all documents will have a table of contents, while only some documents will
have an index.  For many languages (but not for English) both the table of
contents and the index come under the general term \quotation{index}.  For
English readers, a table of contents will normally come at the beginning
(of a document, or possibly in some cases at the beginning of chapters as
well, such as is done in this document), and the index will come at the end.

Either of these imply a particular application of the mechanism for
internal references whose explanation is included in
\in{Section}[sec:references].

\startsection
  [
    reference=sec:content,
    title=Tables of contents,
  ]

\startsubsection
  [title=Overall view of the table of contents]

In the previous chapter we examined the commands that allow the structure
of a document to be established as it has been written.  This section
focuses on the table of contents and the index, which in some way {\em
mirror} the document's structure.  The table of contents is very useful
for getting an idea of the document as a whole (it helps contextualise
information) and for searching for the exact point where a particular
passage might be located.  Books with a very complex structure, with many
sections and subsections with various levels of depth, seem to require a
different kind of table of contents, since a poorly detailed one (perhaps
with only the first two or three levels of sectioning) helps a lot to get
an overall idea of the contents of the document, but is not very useful for
locating a particular passage.  In a very detailed table of contents, on
the other hand, it is easy to miss the forest for the trees and lose the
overall view of the document.  This is why, sometimes, books with a
particularly complex structure include more than one table of contents: one
not too detailed at the beginning showing the main parts, and a more
detailed table of contents at the beginning of each chapter as well as,
perhaps, an index at the end.

These can all be generated by \ConTeXt\ automatically with relative ease.
We can:

\startitemize

  \item Generate a complete or partial table of contents at any point in
    the document.

  \item Decide on the contents of either.

  \item Configure their appearance down to the last detail.

  \item Include hyperlinks in the table of contents that allow us to jump
    directly to the section in question.

\stopitemize

In fact this last utility is included by default in all tables of contents
provided that the interactivity function has been enabled in the document.
This is discussed in \in{Section}[sec:interactivity].

The explanation of tables of contents in the \ConTeXt\ reference manual is,
in my opinion, somewhat confusing, which I think is due to the fact that
too much information is introduced at once. The mechanism for
building \ConTeXt\ tables of contents has many pieces to it, and as a
consequence it is difficult for a text that tries to explain them all at
once to be clear, especially for the reader who is new to the scene.  By
contrast, the explanation in the
\goto{wiki}[url(https://wiki.contextgarden.net/Document_structure_and_headlines/Table_of_contents)],
is essentially limited to examples: very useful for learning {\em tricks}
but inadequate --- I think --- for understanding the mechanism and how it
works.  This is why the strategy I have decided to use to explain things in
this introduction begins by assuming something that is not strictly true
(or not completely true): that there is something in \ConTeXt\ called the
{\em table of contents}.  Starting with this, the {\em normal} commands for
generating the table of contents are explained, and when these commands and
their configuration are well known, I think this is the moment for
introducing --- though at a theoretical rather than more practical level ---
the information on those pieces of the mechanism that have been omitted up
till then.  Knowledge of these additional {\em pieces} allows us to create
much more customised tables of contents than the ones we can call the {\em
normal ones} created with the commands explained up to that point;
however, in most cases we will not need to do this.

\stopsubsection


\startsubsection
  [
    reference=sec:completecontent,
    title=Completely automatic table of contents with a title,
  ]

The basic commands for generating an automatically generated table of
contents (TOC) from the numbered sections of a document ({\tt part}, 
{\tt chapter}, {\tt section}, etc.\null)
are \PlaceMacro{completecontent}\tex{completecontent}
and \PlaceMacro{placecontent}\tex{placecontent}.  The main difference
between the two commands is that the first adds a {\em title} to the TOC;
to do so, immediately before the TOC it inserts an {\em unnumbered chapter}
whose default title is Table of Contents.

Therefore \tex{completecontent}:

\startitemize

  \item Inserts, at the point where it is found, a new unnumbered chapter
    entitled \quotation{Table of Contents}.

    \startSmallPrint

        We recall that in \ConTeXt\ the command used to generate an
        unnumbered section at the same level as chapters, is \tex{title}
        (see \in{Section}[sec:sectiontypes]).  Therefore in reality
        \tex{completecontent} does not insert a {\em Chapter}
        (\tex{chapter}) but a {\em Title} (\tex{title}).  I have not said
        so because I think it may be confusing for the reader to use the
        names of the unnumbered section commands here, since the term {\em
        Title} also has a broader sense, and it is easy for the reader
        not to identify it with the concrete level of sectioning we are
        referring to.

    \stopSmallPrint

  \item This {\em chapter} (actually, \tex{title}) is formatted exactly the
    same as the rest of the unnumbered chapters in the document; by
    default, this includes a page break.

  \item The table of contents is printed immediately after the title.

\stopitemize

Initially the generated TOC is {\em complete}, as we can deduce from the
command name that generates it: (\tex{completecontent}).  But on the one
hand we can limit the level of depth of the TOC as explained in
\in{Section}[sec:placecontent] and, on the other hand, since this command is
{\em sensitive} to the place it is found in the source file (see what is
said further on about \tex{placecontent}), if \tex{completecontent} is not
found at the beginning of the document it is possible that the TOC
generated is not complete.  In some points of the source file it is
even possible that the command is apparently ignored.  If this happens, the
solution is to invoke the command with the \MyKey{criterium=all} option.
Regarding this option, also see \in{Section}[sec:placecontent].

To change the default title assigned to the TOCs we use the
\PlaceMacro{setupheadtext}\tex{setupheadtext} command whose syntax is:

  \type{\setupheadtext[Language][Element=Name]}

where {\tt Language} is optional and refers to the language identifier used
by \ConTeXt\ (see \in{Section}[sec:langdoc]), and {\tt Element} refers to
the element whose name we want to change (\MyKey{content} in the case of
the table of contents) and {\tt Name} is the name or title we want to give
our TOC\null.  For example

  \type{\setupheadtext[en][content=Contents]}

will ensure that the TOC generated by \tex{completecontent} is entitled
\quotation{Contents} instead of \quotation{Table of Contents}.

Moreover, \tex{completecontent} allows the same configuration options as
\tex{placecontent}, for the explanation of which I refer to (the next
section).

\stopsubsection

\startsubsection
  [
    reference=sec:placecontent,
    title=Automatic table of contents without a title,
  ]

The general command for inserting a TOC without a title, generated
automatically from the document's sectioning commands, is
\tex{placecontent}, whose syntax is:

\type{\placecontent [Options]}

In principal, the table of contents will contain absolutely all numbered
sections, although we can limit its level of depth with the
\PlaceMacro{setupcombinedlist}\tex{setupcombinedlist} command (that we will
speak of further on).  So, for example:

\type{\setupcombinedlist[content][list={chapter,section}]}

will limit the contents of the TOC to chapters and sections.

A peculiarity of this command is that it is sensitive to its location in
the source file.  This is very easy to explain with a few examples, but much
more difficult if we want to specify exactly how the command works and
which headings are included in the TOC in each case.  So let's start with
the examples:

\startitemize

  \item \tex{placecontent} placed at the beginning of the document, before
    the first section command (part, chapter or section, according to the
    situation) will generate a complete table of contents.

    \startSmallPrint

        I am not really sure that the table of contents generated by
        default is {\em complete}, I believe it does include enough levels
        of sectioning to be complete in most cases; but I suspect it will
        not go beyond the eighth level of sectioning.  In any case, as
        mentioned above, we can adjust the sectioning level the TOC reaches
        with

        \type{\setupcombinedlist[content][list={chapter, section, subsection, ...}]}

    \stopSmallPrint

  \item By contrast, this same command located inside a part, chapter or
    section will exclusively generate a TOC of the content of that element,
    or in other words chapters, sections and other lower levels of
    sectioning of a specific part, or sections (and other levels) of a
    specific chapter, or subsections of a specific section.

\stopitemize

As for the technical and detailed explanation, in order to understand the
default operation of \tex{placecontent} properly, it is essential to
remember that the various sections are, in fact, {\em environments} for
\ConTeXt\ LMTX that start with \tex{startSectionType} and end with
\tex{stopSectionType} and can be contained within other lower level section
commands.  So, taking that into account, we can say that \tex{placecontent}
generates by default a table of contents that will only include:

\startitemize

  \item Elements that belong to the {\em environment} (section level) where
    the command is placed.  This means that the command when placed in a
    chapter will not include sections or subsections from other chapters.

  \item Elements that have a level lower of sectioning than the level
    corresponding to the point where the command is located.  That is, if
    the command is in a chapter, only sections, subsections and other lower
    levels are included; but if the command is in a section, the index will
    be split to make the TOC of the subsection level.

\stopitemize

Furthermore, for the table of contents to be generated, it is required that
\tex{placecontent} be found {\em before} the first section of the chapter
in which it is located, or before the first subsection of the section in
which it is located, etc.

\startSmallPrint

    I'm not sure I was clear in the explanation above. Perhaps with a
    somewhat more detailed example than the previous ones we can better
    understand what I mean: let's imagine the following structure of a
    document:

\vbox{ \startitemize[packed]

  \item Chapter 1

    \startitemize[packed]

    \item Section 1.1

    \item Section 1.2

      \startitemize[packed]

      \item Subsection 1.2.1

      \item Subsection 1.2.2

      \item Subsection 1.2.3

      \stopitemize

    \item Section 1.3

    \item Section 1.4

    \stopitemize

  \item Chapter 2

  \stopitemize}

    So: \tex{placecontent} placed before Chapter 1 will generate a complete
    table of contents, similar to the one generated
    by \tex{completecontent} but without a title. But if the command is
    placed within Chapter 1 and before section 1.1, the table of contents
    will be only of the chapter; and if it is placed at the beginning of
    section 1.2, the table of contents will be only the content of that
    section. But if the command is placed, for example, between sections
    1.1 and 1.2 it will be ignored. It will also be ignored if it is placed
    at the end of a section, or at the end of the document.

\stopSmallPrint

All of this, of course, refers only to the case where the command does not
include options. In particular, the {\tt criterium} option will alter that
default behaviour.

Of the options allowed by \tex{placecontent} I will only explain two of
them, the most important ones for setting up the TOC, and, moreover, the
only ones that are (partially) documented in the \ConTeXt\ reference
manual. The {\tt criterium} option, which affects the content of the TOC in
relation to the place in the source file where the command is located; and
the {\tt alternative} option, which affects the general layout of the TOC
to be generated.

\stopsubsection

\startsubsection
  [
    reference=sec:criteriumlist,
    title={Elements to incorporate in the TOC: the {\tt criterium} option},
  ]

The default operation of \tex{placecontent} in relation to the position of
the command in the source file has been explained above. The {\tt
criterium} option alters this operation. Among others, it can take the
following values:

\startitemize

  \item {\tt all}: the TOC will be complete, regardless of the place in the
  source file where the command is found.

  \item {\tt previous}: the TOC will only include the section commands (of
  the level we are at) {\em previous} to \tex{placecontent}. This option is
  intended for TOCs that are written at the end of the document or section
  in question.

  \item {\tt part, chapter, section, subsection...}: implies that the TOC
  should be limited to the sectioning level indicated.

  \item {\tt component}: in multifile projects (see
  \in{Section}[sec-projects]), it will generate only the TOC corresponding
  to the {\em component} where the \tex{placecontent} or
  \tex{completecontent} command is found.

\stopitemize

\stopsubsection

\startsubsection
  [
    reference=sec:alternativelist,
    title={Layout of the table of contents: the {\tt alternative} option},
  ]

The {\tt alternative} option controls the overall layout of the table of
contents. Its main values can be see in \in{Table}[tbl:contentalternatives].

\placetable
  [here]
  [tbl:contentalternatives]
  {\tfx Ways of formatting the table of contents}
{\switchtobodyfont[small]
\starttabulate[|c|l|l|]
\HL
\NC {\bf alternative}
\NC {\bf Contents of TOC entries}
\NC {\bf Notes}
\NR
\HL
\NC a
\NC Number -- Title -- Page
\NC One line per entry
\NR
\NC b
\NC Number -- Title -- Spaces -- Page
\NC One line per entry
\NR
\NC c
\NC Number -- Title -- Leader dots -- Page
\NC One line per entry
\NR
\NC d
\NC Number -- Title -- Page
\NC Continuous TOC
\NR
\NC e
\NC Title
\NC Framed
\NR
\NC f
\NC Title
\NC Left aligned,
\NR
\NC\NC\NC right aligned or centred
\NR
\NC g
\NC Title
\NC Centred
\NR
\HL
\stoptabulate}

The first four alternative values provide all the information of each
section (its number, its title and the page number where it begins), and
are therefore suitable for both paper and electronic documents. The last
three alternatives only inform us about the title, so they are only
suitable for electronic documents where it is not necessary to know the
page number where a section begins, provided that the TOC includes a
hyperlink to it, which happens by default in \ConTeXt.

Furthermore, I believe that in order to truly appreciate the differences
between the various alternatives, it is best for the reader to generate a
test document where he or she can analyse them in detail.

\stopsubsection

\startsubsection
  [
    reference=sec:setuplist,
    title=Format of TOC entries,
  ]
  \PlaceMacro{setuplist}

We have seen that the {\tt alternative} option of \tex{placecontent} or
\tex{completecontent} allows us to control the general {\em layout} of the
table of contents, i.e., what information will be shown for each heading,
and whether or not there will be line breaks separating the different
headings. Final adjustments to each TOC entry are made with the
\tex{setuplist} command whose syntax is as follows:

\type{\setuplist[Element][Configuration]}

where {\tt Element} refers to a particular kind of section.  This could be
{\tt part}, {\tt chapter}, {\tt section}, etc.  We can also configure more
than one element at the same time, separating them with commas. {\tt
Configuration} has up to 54 possibilities, many of them, as usual, not
expressly documented; but this does not prevent those that are documented,
or the ones that are not clear enough, from allowing full adjustment of the
TOC\null.

I will now explain the most important options, grouping them according to
their usefulness, but before going into them let us remember that a TOC
entry, depending on the value of the {\tt alternative}, can have up to
three different components: the section number, the title of the section,
and the page number.  The configuration options allow us to configure the
various components globally or separately:

\startitemize

  \item {\em Inclusion (or not) of the different components}: if we have
    chosen an alternative that includes, in addition to the title, the
    section number and the page number (alternatives \quote{{\tt
    a}}, \quote{{\tt b}}, \quote{{\tt c}} or \quote{{\tt d}}), the
    options {\tt headnumber=no} or {\tt pagenumber=no} mean that for the
    specific level we are configuring, the section number ({\tt
    headnumber}) or the page number ({\tt pagenumber}) is not
    displayed.

  \item {\em Colour and style}: we already know that the entry that
    generates a specific section in the TOC may have (depending on the
    alternative) up to three different components: section number, title
    and page number.  We can jointly indicate the style and the colour for
    the three components using the {\tt style} and {\tt color} options, or
    do it individually for each component by means of {\tt numberstyle},
    {\tt textstyle} or {\tt pagestyle} (for the style) and {\tt numbercolor},
    {\tt textcolor} or {\tt pagecolor} for the colour.

    To control the appearance of each entry, in addition to the style
    itself, we can apply some command to the whole entry or to one of its
    different elements.  For this there are the {\tt command}, {\tt
    numbercommand}, {\tt pagecommand} and {\tt textcommand} options.
    The command indicated here can be a standard \ConTeXt\ command or a
    command of our own creation.  The section number, title text and page
    number will be passed as arguments to the {\tt command} option, while
    the section title will be passed as an argument to {\tt textcommand}
    and page number to {\tt pagecommand}.  So, for example, the following
    instruction will ensure that section titles are written in (fake) small
    caps:

    \starttyping

    \setuplist[section][textcommand=\Cap]

    \stoptyping

  \item {\em Separation of the other TOC elements}: the {\tt before} and
    {\tt after} options allow us to indicate the commands that will be
    executed before ({\tt before}) and after ({\tt after}) typesetting the
    TOC entry.  Normally these commands are used to set either the spacing
    or some separating element between the previous and subsequent entries.

  \item {\em Indenting an element}: set with the {\tt margin} option which
    allows us to set the amount of left indentation that the entries of the
    level we are configuring will have.

  \item {\em Hyperlinks embedded in the TOC}: by default the index entries
    include a hyperlink to the document page where the section in question
    begins.  Using the {\tt interaction} option we can disable this
    function ({\tt interaction=no}) or limit the part of the index entry
    where the hyperlink will be, which can be the section number ({\tt
    interaction=number} or {\tt interaction=sectionnumber}), the
    section title ({\tt interaction=text} or {\tt interaction=title}) or
    the page number ({\tt interaction=page} or {\tt interaction=pagenumber}).

\need3\baselineskip

  \item {\em Other aspects}:

  %TODO: I didn't see width do anything in a quick test.
  % And what does symbol actually do?  Wiki is silent in March 2026.

    \startitemize

      \item {\tt width}: specifies the separation distance between the
        number and title of the section.  It can be a dimension, or the
        keyword {\tt fit} that sets the exact width of the section number.

% TODO: an example of what symbol=one (or two or three) does would be illuminating.

      \item {\tt symbol}: allows the section number to be replaced by a
        {\em symbol}.  Three possible values are supported: {\tt one}, {\tt
        two} and {\tt three}.  The value {\tt none} for this option
        removes the section number from the TOC\null.

      \item {\tt numberalign}: indicates the alignment of numbering
        elements; it can be left, right, middle, flushright, flushleft.

    \stopitemize
    
\stopitemize

Among the multiple configuration options of the TOC, there are none that
allows us to directly control the interline spacing.  This will be, by
default, the one that applies to the document as a whole.  Often, however,
it is preferable that lines in the TOC are slightly {\em tighter} than the
rest of the document.  To achieve this we should enclose the command that
generates the table of contents (\tex{placecontent} or
\tex{completecontent}) within of a group where a different interline
spacing is established.  For example:

\starttyping
\start
  \setupinterlinespace[small]
  \placecontent
\stop
\stoptyping

\stopsubsection


\startsubsection
  [
    reference=sec:manual adjustments,
    title=Manual adjustments to the table of contents,
  ]

We have already explained the two fundamental commands for generating
tables of contents (\tex{placecontent} and \tex{completecontent}), as well
as their options.  With these two commands, TOCs are automatically
generated, constructed from the existing numbered sections in the document
(or in the block or segment of the document to which the table of contents
refers). I will now explain certain {\em settings} that we can make so that
the content of the TOC is not so {\em automatic}.  This implies:

\startitemize

  \item The possibility of also including some unnumbered section titles in
    the TOC\null.

  \item The possibility of manually sending a particular entry to the TOC
    that does not correspond to the presence of a numbered section.

  \item The possibility of excluding a particular numbered section from the
    TOC\null.

  \item The possibility that the title for a particular section reflected
    in the TOC does not coincide exactly with the title included in the
    body of the document.

\stopitemize

\stopsubsection

% TODO: The size and boldness of the subsubsection titles make them look
% non-subservient to the subsections.  Anyone reading this agree?

\startsubsubsection
  [
    reference=sec:toc with unnumbered secs,
    title=Including unnumbered sections in the TOC,
  ]

The mechanism by which \ConTeXt\ builds the TOC means that all numbered
sections are automatically included, which, as I have already said (see
\in{Section}[sec:title parts]) depends on the two ({\tt number} and {\tt
incrementnumber}) options that we can change with \tex{setuphead} for each
kind of section.  It was also explained there that a section type where {\tt
incrementnumber=yes} and {\tt number=no} would be an internally but not
externally numbered section.

Therefore, if we want a particular unnumbered section type --- for example,
{\tt title} --- to be included in the TOC, we must change the value of the
{\tt incrementnumber} option for that section type, setting it to {\tt yes}
and then include that section type among those to be displayed in the TOC,
which is done, as explained above, with \tex{setupcombinedlist}:

\starttyping
  \setuphead
    [title]
    [incrementnumber=yes]

 \setupcombinedlist
   [content]
   [list={chapter, title, section, subsection, subsubsection}]
\stoptyping

We can then, if we wish, format this entry using \tex{setuplist} in exactly
the same way as any of the others; for example:

\type{\setuplist[title][style=bold]}

{\bf Note:} The procedure just explained will include all instances in our
document of the unnumbered section type concerned (in our example the {\tt
title} type sections).  If we only wish to include a particular occurrence
of that section type in the TOC, it is preferable to do so by the procedure
explained below.

\stopsubsubsection

\startsubsubsection
  [
    reference=sec:manualtoc,
    title=Manually adding entries to the TOC,
  ]

From any point in the source file we can submit either an entry (simulating
the existence of a section that does not really exist) or a command to the
table of contents.

To send an entry that simulates the existence of a section that does not
really exist, use the \PlaceMacro{writetolist}\tex{writetolist} command,
whose syntax is:

\type{\writetolist[SectionType][Options]{Number}{Text}}

in which:

\startitemize

  \item The first argument indicates the level that this section entry must
    have in the TOC: {\tt chapter}, {\tt section}, {\tt subsection}, etc.

% I am not sure what is meant here.  If the manually submitted entry is
% omitted, then what are we talking about?

  \item The second argument, which is optional, allows this entry to be
    configured in a particular way.  If the manually submitted entry is
    omitted, it will be formatted like all the entries of the level
    indicated with the first argument; although, I must point out that in
    my tests I have not managed to make it work.

    \startSmallPrint

        Both in the official list of \ConTeXt\ commands (see
        \in{Section}[sec:qrc-setup-en]) and in the wiki we are told that
        this argument allows the \Doubt same values as \tex{setuplist}
        which is the command that allows us to format the different TOC
        entries.  But, I insist, in my tests I have not managed to change
        the appearance of the TOC entry sent manually in any way.

    \stopSmallPrint

  \item The third argument is supposed to reflect the numbering that the
    element sent to the \Doubt TOC has, but I couldn't get this to work in
    my tests either.

  \item The last argument includes the text to be sent to the TOC\null.

\stopitemize

This is useful, for example, if we want to add a particular unnumbered
section to the TOC, but only thatone.  In \in{Section}[sec:toc with
unnumbered secs] it explains how to add an entire category of
unnumbered sections to the table of contents; but if we only
want to add a particular occurrence of a section type to it, it is more
convenient to use the \tex{writetolist} command.  And so, for example, if we
want the section of our document containing the bibliography to be an
unnumbered section, but still included in the TOC, we would write:

\starttyping
\subject{Bibliography}
\writetolist[section]{}{Bibliography}
\stoptyping

See how we are using the unnumbered version of {\tt section}, which is {\tt
subject}, for the section but we are adding it to the index, manually,
as if it were a numbered section ({\tt section}).

Another command intended to influence the table of contents manually is
\PlaceMacro{writebetweenlist}\tex{writebetweenlist} which is used to add
not an entry itself, but a {\em command} to the table of contents, from a
particular point in the document.  For example, if we want to include a line
between two items in the TOC, we could write the following at any point in
the document located between the two sections concerned:

\type{\writebetweenlist[section]{\hrule}}

\stopsubsubsection

\startsubsubsection
    [title=Exclude from the TOC a specific section belonging to a type of
        section that is otherwise included]

% Original English title, the above from another translation
%  [title=Exclude a particular section from the TOC belonging to a section
%      type that is included in the TOC]

The table of contents is constructed from {\em section types} established,
as we already know, by the {\tt list} option of \tex{setupcombinedlist};
consequently, if a certain {\em section type} appears in the TOC, there is
no way of excluding a particular section (of that type) we don't want in
the TOC for whatever reason(s).

Normally, if we don't want a section to appear there, what we would do is
to use its {\em unnumbered equivalent} meaning, for example, {\tt title}
instead of {\tt chapter}, {\tt subject} instead of {\tt section}, etc.
These sections are not sent to the TOC, and neither are they numbered.

However, if for any reason we want a certain section to be numbered but not
appear in the table of contents, even if other types of this kind do, we
can use a {\em trick} which consists of creating a new section type that is
a clone of the section in question.  For example:

\starttyping
\definehead[MySubsection][subsection]
\section{First section}
\subsection{First subsection}
\MySubsection{Second subsection}
\subsection{Third subsection}
\stoptyping

This will ensure that when inserting a section type {\tt MySubsection} the
subsection counter will increase, since this section is a {\em clone} of
the subsections, but the TOC will not include the title or number of such a
section, since by default it does not include {\tt MySubsection} types.

\stopsubsubsection


\startsubsubsection
  [title=Section title text which differs in the TOC from the title in the
      body of the document]

% TODO: it seems \nolists no longer exists (so says the wiki in 2026/03),
% so this section was rewritten. Is there something else which should be
% added instead?


If we do not want the title of a particular section included in the TOC to
be identical to the one displayed in the body of the document, the
following procedure is available to us.  Create the section not with
traditional syntax (\type{SectionType{Title}}) but
with \tex{SectionType[Options]}, or with \tex{startSectionType[Options]},
and assign the text we want to be written in the TOC to the {\tt list}
option (see \in{Section}[sec:sectionsyntax]).  

For example, if a section title is very long, and a truncated version is
desired for the TOC, this could be achieved as follows:

\starttyping
\startsection
  [
    reference=sec:investigations,
    title={Discovering why the sea is boiling hot, how pigs gained the
        power of flight, and whether snarks can, in fact, be defeated with
        vorpal blades},
    list={Discovering why the sea is boiling hot, etc.},
  ]
\stoptyping

% If we do not want the title of a particular section included in the TOC to
% be identical to the one displayed in the body of the document, we have two
% procedures available to us:
% 
% \startitemize
% 
%   \item Create the section not with traditional syntax
%     (\type{SectionType{Title}}) but with \tex{SectionType[Options]}, or
%     with \tex{startSectionType[Options]}, and assign the text we want to be
%     written in the TOC to the {\tt list} option (see
%     \in{Section}[sec:sectionsyntax]).
% 
%   \item When writing the title of the section in question in the body of
%     the document, use the \tex{nolist} command: this command causes the
%     text it takes as an argument to be replaced in the TOC by an
%     ellipsis. For example:
% 
%     \starttyping
% \chapter
%   [title={An \nolist{approximate and slightly repetitive}
%            introduction to the reality of the obvious}]
%     \stoptyping
% 
%     would typeset as the chapter title in the body of the document,
%     \quotation{An approximate and slightly repetitive introduction to the
%     reality of the obvious}, but would send the following text to the TOC
%     \quotation{An ... introduction to the reality of the obvious}.
% 
%     \startSmallPrint
% 
%         {\bf Attention:} What I have just pointed out about the \tex{nolist}
%         command is stated in both the \ConTeXt\ reference manual and the
%         \goto{wiki}[url(https://wiki.contextgarden.net/Command/nolist)]. For
%         me, however, it produces a compiling error, telling me that the
%         \tex{nolist} command is undefined.
% 
%     \stopSmallPrint
% 
% \stopitemize

\stopsubsubsection

\stopsubsection

\stopsection


\startsection
  [
    reference=sec:lists,
    title={Lists, combined lists and tables of contents based on lists},
  ]

Internally, for \ConTeXt, a table of contents is nothing more than a {\em
combined list}, which, in turn, as its name suggests, consists of a
combination of simple lists.  Therefore the basic notion from which
\ConTeXt\ builds the table of contents is that of a list.  Several lists
are combined to form a table of contents.  By default, \ConTeXt\ contains a
predefined combined list called \MyKey{content} and this is what the
commands examined so far work with: \tex{placecontent} and
\tex{completecontent}.

\startsubsection
    [title=Lists in \ConTeXt]

% TODO: what about unnumbered things?

In \ConTeXt, a {\em list} is a range of numbered elements about which we
need to remember three things:

\startitemize[n,packed]

  \item The number.

  \item The name or title.

  \item The page where it is found.

\stopitemize

This happens not only with numbered sections, but also with other elements
of the document such as images, tables, etc.  In general, it happens for
those elements for which there is a command whose name begins
with \tex{place}, such as \tex{placetable}, \tex{placefigure}, etc.

In all of these cases, \ConTeXt\ automatically generates a list of the that
collects the different occurrences of the element type in question, its
number, title and page.  Thus, for example, there is a list of chapters,
called {\tt chapter}, another of sections, called {\tt section}; but also
another of tables (called {\tt table}) and another of figures (called {\tt
figure}).  Lists generated automatically by \ConTeXt\ are always named
the same as the type of element they store.

A list will also be automatically generated if we create, for example, a
new type of numbered section: creating a new numbered section will
implicitly also create the list that stores them.  And if, for a
section that is not numbered by default, we set the option {\tt
incrementnumber=yes} (making it a numbered section), we will also be
implicitly creating a list with that name.

Along with the implicit lists (automatically defined by \ConTeXt) we can
create our own lists with \tex{definelist}, whose syntax is:

\PlaceMacro{definelist}\type{\definelist[ListName][Configuration]}

\need3\baselineskip

Items on the list are added:

\startitemize

  \item In lists predefined by \ConTeXt, or created by it as a result of
    creating a new floating object (see \in{Section}[sec:definefloat]), the
    corresponding entry will be automatically inserted into its list
    whenever an item is added to the document, either by using a sectioning
    command or by using the \tex{placeWhatever} command for other types of
    lists.  For example: \tex{placefigure} will insert an image into the
    document, but it will also insert the corresponding entry into the list.

  \item In any type of list, manually with \tex{writetolist[ListName]},
    as already explained in \in{Subsection}[sec:manualtoc] of
    \in{Section}[sec:manual adjustments].  The \tex{writebetweenlist}
    command is also available, explained in that same section.

\stopitemize

Once a list has been created and all its items included, the three
basic commands related to it are \tex{setuplist},
\PlaceMacro{placelist}\tex{placelist} and
\PlaceMacro{completelist}\tex{completelist}.  The first allows us to
configure what the list looks like; the latter two insert the list in
question at the point in the document where they are found.  The difference
between \tex{placelist} and \tex{completelist} is similar to the difference
between \tex{placecontent} and \tex{completecontent} (see sections
\in{}[sec:completecontent] and \in{}[sec:placecontent]).

So, for example:

\type{\placelist[section]}

will insert a list of the sections, including a hyperlink to them if the
document's interactivity is enabled and if, in \tex{setuplist}, we have not
set {\tt interaction=no}.  A list of sections is not exactly the same as a
table of contents based on sections: the idea of a table of contents
usually includes the lower levels as well (subsections, subsubsections,
etc.).  But a list of sections will include only the sections themselves.

The syntax of these commands is:

\type{\placelist[ListName][Options]}

\type{\setuplist[ListName][Configuration]}

The \tex{setuplist} options have already been explained in
\in{Section}[sec:setuplist], and the options for \tex{placelist} are the
same as for \tex{placecontent} (see \in{Section}[sec:placecontent]).

\stopsubsection


\startsubsection
  [
    reference=sec:variouslists,
    title={Lists or indices of images, tables and other items},
  ]

From what has been said so far, it can be seen that, since \ConTeXt\
automatically creates a list of images placed in a document with the
\tex{placefigure} command, generating a list or index of figures at a
particular point in our document is as simple as using the
\tex{placelist[figure]} command.  If we want to generate a list with a
title (similar to what we get with \tex{completecontent}) we can do it with
\tex{completelist[figure]}.  Similarly, we can use \tex{placelist}
and \tex{completelist} with the other four predefined kinds of floating
objects in \ConTeXt: tables (\MyKey{table}), graphics (\MyKey{graphic}),
{\em intermezzos} (\MyKey{intermezzo}) and chemical formulas
(\MyKey{chemical})

For each of the predefined list types, \ConTeXt\ already includes a command
that generates them without a title
(\PlaceMacro{placelistoffigures}\tex{placelistoffigures},
\PlaceMacro{placelistoftables}\tex{placelistoftables},
\PlaceMacro{placelistofgraphics}\tex{placelistofgraphics},
\PlaceMacro{placelistofintermezzi}\tex{placelistofintermezzi} and
\PlaceMacro{placelistofchemicals}\tex{placelistofchemicals}), and another
that generates them with a title
(\PlaceMacro{completelistoffigures}\tex{completelistoffigures},
\PlaceMacro{completelistoftables}\tex{completelistoftables},
\PlaceMacro{completelistofgraphics}\tex{completelistofgraphics},
\PlaceMacro{completelistofintermezzi}\tex{completelistofintermezzi} and
\PlaceMacro{completelistofchemicals}\tex{completelistofchemicals}), in a
similar way to \tex{completecontent}.
{\em Note}: while the common English plural of \quotation{intermezzo} may
be \quotation{intermezzos}, the \ConTeXt\ \tex{placelistof...}
and \tex{completelistof...} commands use {\em intermezzi}.

In the same way, for floating objects we ourselves have created (see
\in{Section}[sec:definefloat]) the \tex{placelistof<FloatName>} and
\tex{completelistof<FloatName>} commands will be automatically created.

For lists we create ourselves with \tex{definelist} we can create an index
with \tex{placelist[ListName]} or \tex{completelist[ListName]}.

\stopsubsection


\startsubsection
  [title=Combined lists]

A combined list is, as its name suggests, a list that combines items from
different previously defined lists.  By default, \ConTeXt\ defines a
combined list of content indices by default named \MyKey{content}, but we
can create other combined lists
with \PlaceMacro{definecombinedlist}\tex{definecombinedlist}, whose syntax
is:

\tex{definecombinedlist[Name][Lists][Options]}

where:

\startitemize

  \item {\tt Name} is the name the new combined list will have.

  \item {\tt Lists} refers to the names of lists to be combined, separated
    by commas.

  \item {\tt Options} is the selection of configuration options for the
    list.  These options can be indicated when the list is defined, or,
    probably preferably, when the list is invoked.  The main options (which
    have already been explained) are {\tt criterium}
    (\in{Subsection}[sec:criteriumlist] of \in{Section} [sec:placecontent])
    and {\tt alternative} (in \in{Subsection}[sec:alternativelist] in the
    same section).

\stopitemize

A side effect of creating a combined list with \tex{definecombinedlist} is
that a command called \tex{placeListName} is also created, which is used to
invoke the list, i.e., to include it in the output file.  For example:

\tex{definecombinedlist[TOC]}

will create the command \tex{placeTOC}; and:

\tex{definecombinedlist[content]}

will create the command \tex{placecontent}.

But wait, \quotation{\tex{placecontent}}?  Isn't this the command that is
used to create a {\em normal} table of contents?  Indeed!  This means that
the standard table of contents is actually created by \ConTeXt\ by means of
the following command:

\starttyping
\definecombinedlist
  [content]
  [part, chapter, section, subsection,
    subsubsection, subsubsubsection,
    subsubsubsubsection]
\stoptyping

Once our combined list is defined, we can configure it (or reconfigure it)
with \tex{setupcombinedlist}, which allows the use of the already explained
options {\tt criterium} (see \in{Subsection}[sec:criteriumlist]) and {\tt
alternative} (see \in{Subsection}[sec:alternativelist]), as well as the
{\tt list} option to {\em change} the lists included in the combined list.

\startSmallPrint

    The official list of \ConTeXt\ commands (see
    \in{Section}[sec:qrc-setup-en]) does not mention the {\tt list} option
    among the options allowed for \tex{setupcombinedlist}, but it is used in
    several examples of the use of this command in the wiki (which, moreover,
    does not mention it in the page devoted to this command either). I have
    also checked that the option works.

\stopSmallPrint

\stopsubsection

\stopsection


% TODO: At least one day, without this \need the Section title was left all
% by itself.  Presumably the \subsection "\need" and the \section "\need"
% have to be adjusted.  Or maybe just always use a \need before a \section
% which is immediately followed by a \subsection.
\need 2.7in

\startsection
  [title=Indices]

\startsubsection
  [title=Generating an index]

A {\em subject index}, usually located at the end of a document, consists
of a list of significant terms, indicating the pages where each such term
can be found in the document.

When books were typeset by hand, generating a subject index was a complex
task, as well as a tedious one.  A change in the pagination could affect
many (if not all) of the entries in the index.  Therefore, indices were not
very common.  Today, the existence of computer typesetting programs means
that, while the task is likely to continue being tedious, it is no longer
so complex, given that it is easy for a computer program to keep the data
associated with each index entry up to date.

To generate a subject index we need to:

\startitemize[n]

  \item Determine which words, terms or concepts are to be part of it.
    This is a task that is really up to the author (or another person
    equally familiar with the document's subject matter).

  \item Find at which point(s) in the document each entry in the future
    index appears.  Although, to be precise, rather than just {\em finding}
    the places in the source file where the concept or issue is discussed,
    when we work with \ConTeXt\ we {\em mark} those spots, by inserting
    a command that will then serve to generate the index automatically.
    This is the tedious part.

  \item Generate and format the index by placing a command at the point of
    our choice in the document.  The latter is quite simple with
    \ConTeXt\ and requires only one command: \tex{placeindex}.

\stopitemize

% TODO: this subsection title wasn't really clear to me.
% I made a change, but I think a better title yet is somewhere out there.
%\subsubsection{The prior definition of the entries in the index and the
%marking of the points in the source file that refer to them}
\subsubsection{Defining the entries in the index and 
marking points in the source file to which they refer}

The fundamental work is in the second step.  It is true that computer
systems also facilitate this, in the sense that we can do a global text
search to locate the places in the source file where a specific subject is
discussed.  But we should also not blindly rely on such text searches: a good
subject index must be able to detect every spot where a particular subject
is being discussed, even if this is done without using the {\em standard}
term to refer to it.  Further, in general there will be places where a
particular term is used, but not in an illustrative way that deserves a
mention in the index.

To {\em mark} an actual point in the source file, associating it with a
word, term or idea that will appear in the index, we use the
\PlaceMacro{index}\tex{index} command whose syntax is as follows:

\type{\index[Alphabetical]{Index entry}}

where {\tt Alphabetical} is an optional argument that is used to indicate
an alternative text to that of the index entry itself in order to sort it
alphabetically, and {\tt Index entry} is the text that will appear in the
index, associated with this mark.  We can also apply the formatting features
that we wish to use, and if reserved characters appear in the text, they
must be written in the usual way in \ConTeXt.

\startSmallPrint

% TODO: huh, in my tests with LMTX, \index{\TeX} ignores the \ and puts
% the entry under 't'.  So this example is bogus.

%  The possibility of alphabetising an index entry in a way different from
%  how it is actually written is very useful.  Think, for example, of this
%  document, where I might want to generate an index entry for all
%  references to the \tex{TeX} command.  The command \type{\index{\TeX}}
%  will list \tex{TeX} not by the \quote{t} of \quotation{TeX}, but among
%  the symbols, since the term sent to the index begins with a backslash.
%  If we want \tex{TeX} to be listed under \quote{t}, we can
%  write \type{\index[tex]{\TeX}}.

The possibility of alphabetising an index entry in a way different from
how it is actually written is very useful.  Think, for example, of this
document, where it might be useful to generate an index entry for all
references to the \tex{TeX} command.  You might just try
\starttyping
\index{\TeX}
\stoptyping but that will create an index entry (under \quote{t})
for \quotation{\TeX}, rather than for \quotation{\type{\TeX}}.  If you
actually want to see \type{\TeX} in the index, then you might then try
\starttyping
\index{\backslash TeX}
\stoptyping but that puts the index entry under \quote{b}.  (\ConTeXt\ LMTX
ignores any leading \quote{\backslash} when choosing the letter under which
the index entry will be listed.)  To get \type{\TeX} under \quote{t}, you
can use
\starttyping
\index[tex]{\backslash TeX}
\stoptyping


\stopSmallPrint


\iffalse
% Sample test left here for future experiments.
\starttext

\input knuth

\index[dog]{dog+Labrador}
\index[dog]{dog+Berner}
\index[dog]{dog+Fuzzy}
\index[dog]{Terrier}
\index{dog}
\index[cat]{mountain lion}
\index{*mountain lion}
\index[texxy]{\backslash TeXXy}
\index{\backslash TeXXXy}
\index{\TeX}
\index{\backslash TeX}
\index[texx]{\backslash ppTeX}
\index[gravenhage]{’s Gravenhage} 
\placeindex
\stoptext
\fi

The {\em index entries} will be whatever we choose.  For a subject index to
be really useful we have to work a little harder when considering which
concepts the reader of a document is most likely to look for; so, for
example, it may be better to define an entry as \quotation{disease,
Hodgkin's} than defining it as \quotation{Hodgkin's disease}, since the
more inclusive term is
\quotation{disease}.

\startSmallPrint

    % TODO: The (long) second sentence is in the Spanish ANSS, but not in
    % the original English translation.

    By convention, entries in a subject index are always written in lower
    case, unless they are proper names.
    Martínez de Sousa also suggests using capital letters when a comma has
    been used to place the most significant word before it, and this word
    begins with a capital letter; as in the previous example:
    \quotation{Hodgkin, Disease of} (in Orthography and Orthotypography of Current
    Spanish, 2nd ed., Trea editions 2008, p. 506).

\stopSmallPrint

If the index has several levels of depth (up to three are supported) to
associate a particular index entry with a specific level the \quote{+}
character is used, as follows:

\starttyping
\index{Entry 1+Entry 2}
\index{Entry 1+Entry 2+Entry 3}
\stoptyping

In the first case we defined a second level entry called \MyKey{Entry 2} that
will be a sub-entry of \MyKey{Entry 1}.  In the second case we defined a
third level entry called \MyKey{Entry 3} that will be a sub-entry
of \MyKey{Entry 2}, which in turn is a sub-entry of \MyKey{Entry 1}.  For
example: 

\starttyping
My \index{dog}dog is a \index{dog+greyhound}greyhound called Rocket.
He does not like \index{cat+stray}stray cats.
\stoptyping

\need 3\baselineskip

It is worth noting some details of the above:

\startitemize

  \item The \tex{index} command is usually placed {\em before} the word it
    is associated with and is normally not separated from it by a a blank
    space.  This is to ensure that the command is on the exact same page as
    the word it is linked to:

    \startitemize

      \item If there were a space separating them, there could be the
        possibility that \ConTeXt\ would choose just that space for a line
        break which could also end up being a page break, in which case the
        \tex{index} command would be on one page and the word it is
        associated with on the next page.

      \item If the command were to come {\em after} the word, it would be
        possible for this word to be broken into syllables by the
        hyphenation process, and a line break inserted between two of its
        syllables could also be a page break, in which case the \tex{index}
        entry would refer to page with the end of the word, rather than the
        page with the beginning of the word.

    \stopitemize

  \item Second level terms are introduced in the second and third
    uses of the \tex{index} command.

% TODO: originally this had "cats" where I now have "stray".  In Spanish
% "stray" comes after "cat", and so the example was a little better there,
% contrasting "cat" and "cats".
% A better English language example is badly needed.

  \item Also note how, in the third use of the \tex{index} command,
    although the word that appears in the text is \quotation{stray}, the
    term that will be inserted into the index is \quotation{cat}.

  \item Finally: see how three entries for the subject index have been
    written in just two lines.  I said before that marking the precise
    places in the source file is tedious.  I will now add that marking too
    many of them is counter-productive.  Too extensive an index is by no
    means preferable to a more concise one in which all the information is
    relevant.  That is why I said before that deciding which words will
    appear in the index should be the result of a conscious decision by the
    author.

\stopitemize

If we want our index to be truly useful, terms that are used as synonyms
must be grouped in the index under one head term.  But since it is possible
for the reader to search the index for information by any of the other head
terms, it is common for the index to contain entries that refer to other
entries.  For example, in the subject index of a civil law manual we could
easily find something like:

% TODO: the original amount of whitespace around this was too much.
% The before and after \vskips are horrible hacks.  Do this the right way.

\startframedtext[frame=off,after=\vskip-15pt,before={\null\vskip-25pt}]
  contractual invalidity\\
  \qquad see {\em nullity}.
\stopframedtext

We achieve this not with the \tex{index} command but with
\PlaceMacro{seeindex}\tex{seeindex} whose format is:

\type{\seeindex [Alphabetical] {Entry1} {Entry2}}

where {\tt Entry1} is the index entry that will refer to the other, and
{\tt Entry2} is the reference target.  For our previous example we would write:

\starttyping
\seeindex{contractual invalidity}{nullity}
\stoptyping

In \tex{seeindex} we can also use the \quote{+} sign to indicate sub-levels
for either of its two arguments in curly brackets.

\subsubsection{Generating the final index}

Once we have marked all the entries for the index in our source file, the
actual generation of the index is carried out using the
\PlaceMacro{placeindex}\tex{placeindex} or
\PlaceMacro{completindex}\tex{completindex} commands.  These two commands
(conceptually) scan the source file for the \tex{index} commands, and
generate a list of all the entries that the index should have, associating
a term with the page number corresponding to where it found the \tex{index}
command.  Then they alphabetically order the list of terms that appear in
the index and merge cases where the same term appears more than once, and
finally, they insert the correctly formatted result in the final document.

The difference between \tex{placeindex} and \tex{completeindex} is similar
to the difference between \tex{content} and \tex{completecontent} (see
\in{Section}[sec:completecontent]): \tex{placeindex} is limited to
generating the index and inserting it, while \tex{completeindex}
inserts a new chapter in the final document, called \quotation{Index} by
default, inside which the index will be typeset.

\stopsubsection

\startsubsection
  [title=Formatting the subject index]
  \PlaceMacro{setupregister}

A subject index is a particular application of a more general structure
that \ConTeXt\ calls a \quotation{\em register}; therefore the index is
formatted with the command:

\type{\setupregister[index][Configuration]}

With this command we can:

\startitemize

    \starthead Determine what the index will look like with its different
        elements.  Namely:
    \stophead   %\head only doesn't work wikth LMTX

% TODO: Not really sure what the next sentence is really supposed to be.
%       I was not enlightened by the original Spanish.  What does "and
%       command" mean??

    \startitemize
        
      \item The index headings, which are usually letters of the alphabet,
        can be modified with style, colour and command.  By
        default these letters are in lower case.  With {\tt alternative=A}
        we can set them to be in upper case.

      \item The entries themselves, and their page number.  The appearance
        depends on the {\tt textstyle, textcolor, textcommand} and {\tt
        deeptextcommand} options for the actual entry, and {\tt
        pagestyle, pagecolor} and {\tt pagecommand}, for the page
        number.  With {\tt pagenumber=no} we can also generate a subject
        index without page numbers (although I don't know if this could be
        useful).

      \item The {\tt distance} option measures the width of separation
        between the name of an entry and the page numbers; however, it also
        measures the amount of indentation for subentries.

    \stopitemize

    The names of the {\tt style}, {\tt textstyle}, {\tt pagestyle}, {\tt
    color}, {\tt textcolor}, and {\tt pagecolor} options are clear
    enough to tell us what each one does, I think.  For {\tt command}, {\tt
    pagecommand}, {\tt textcommand} and {\tt deeptextcommand}, I refer
    to the explanation for similarly named options
    in \in{Section}[sec:titlestyle], regarding the configuration of section
    commands.

  \item Set the general appearance of the index, which includes,
    among others, the commands to execute before ({\tt before}) or
    after ({\tt after}) the index, the number of columns it needs to
    have ({\tt n}), whether the columns should be equal or not ({\tt
    balance}), the alignment of entries ({\tt align}), etc.

\stopitemize

\stopsubsection


\startsubsection
    [title=Creating other indices]
  \PlaceMacro{defineregister}\PlaceMacro{setupregister}

I have explained the subject index as if only one such index would be
possible in a document; the truth is that documents can have as many
indices as desired.  There could be an index of personal names (an {\em
onomastic index}), which collects the names of people mentioned in the
document and indicates the place where they are cited; this is a specific
kind of index.  In a legal text we could also create a special index for
mentions of the Civil Code; or, in a document like the present one, an
index of macros explained within, etc.

To create an additional index in our document we use the
\tex{defineregister} command, whose syntax is:

\type{\defineregister [IndexName] [Configuration]}

where {\tt IndexName} is the name the new index will have, and {\tt
Configuration} controls how it works.  It is also possible to configure the
index later on by means of:

\type{\setupregister [IndexName] [Configuration]}

Once a newly named index {\tt IndexName} has been created we will have the
\tex{IndexName} command at our disposal to mark the entries that this index
will have in a similar way to the way entries are marked with \tex{index}.
The \tex{seeIndexName} command also lets us create entries that refer to
other entries.


\need 4\baselineskip

For example: we could create an index of \ConTeXt\ commands in this
document with the command:

\type{\defineregister[macro]}

that would create the \tex{macro} command.  This lets us mark all the
references to \ConTeXt\ commands as an index entry, and then generate the
index with \tex{placemacro} or \tex{completemacro}.

\startSmallPrint

    Creating a new index enables the \tex{IndexName} command to mark its
    entries, and creates corresponding \tex{placeIndexName}
    and \tex{completeIndexName} commands for generating the index.  But
    these latter two commands are actually abbreviations of two more
    general commands applied to the index in question.
    Thus, \tex{placeIndexName} is equivalent to
    \tex{placeregister[IndexName]} and \tex{completeIndexName} is equivalent
    to \tex{completeregister[IndexName]}.

\stopSmallPrint

\stopsubsection

\stopsection

\stopchapter

\stopcomponent

%%% Local Variables:
%%% mode: ConTeXt
%%% eval: (auto-fill-mode 1)
%%% TeX-master: "../introCTX_eng.tex"
%%% coding: utf-8-unix
%%% End:
%%% vim:set filetype=context tw=72 : %%%
