%%% File:      b12_SpecialParas.tex
%%% Author:    Joaquín Ataz-López et al
%%% Begun:     July 2020
%%% Concluded: August 2020
%%% Contents:  This chapter is a big final mishmash.  Everything
%%%            that could not be clearly located elsewhere, is
%%%            here.  I decided on the final structure when
%%%            I was writing this chapter, since when starting
%%%            to deal with certain material I became aware
%%%            that it could be placed elsewhere.
%%%
%%% Edited: Emacs + AUCTeX - And at times vim + context-plugin
%%%

\environment introCTX_env.tex

\startcomponent b12_SpecialParas.tex

\startchapter
    [title={Special constructions and paragraphs}]

\TocChap

\startsection
    [title={Footnotes and endnotes}]

Notes are \quotation{secondary textual elements employed for various
purposes, such as clarifying or extending the main text, providing the
bibliographic reference for the sources, including citations, referring to
other documents or stating the meaning of the text} [{\em Libro de Estilo
de la Lengua española} (Spanish Language Style Guide), p.~195].  They are
particularly important in texts of an academic nature.  They can be placed
at different points on the page or in the document.  Today, the most
widespread ones are those located at the foot of the page (called,
therefore, footnotes); sometimes they are also located in one of the
margins (margin notes), at the end of each chapter or section, or at the
end of the document (endnotes).  In particularly complex documents, there
may also be different {\em series} of notes: author's notes, translator's
notes, updates, etc.  In particular, in critical editions the note apparatus
can become rather complex and only a few typesetting systems are capable of
supporting it; \ConTeXt\ is one of these.  There are numerous commands
available to establish and configure different types of notes.

To explain this, it is useful to start by pointing out the various elements
that can be involved in a note:

\startitemize

  \item The {\em mark} or note {\em anchor}: the sign placed in the body of
    the text to indicate that there is a note linked to it.  Not all types
    of notes have an {\em anchor} associated with them, but when there is
    one, this {\em anchor} appears in two places: at the point in the main
    text to which the note refers, and at the beginning of the note text
    itself.  The presence of the same reference mark in both places is what
    allows the note to be associated with the main text.

  \item The note {\em ID or identifier}: the letter, number or symbol that
    identifies the note and distinguishes it from other notes.  Some notes,
    for example margin notes, can lack an ID\null.  When this is not the case,
    the ID normally coincides with the {\em anchor}.

    \startSmallPrint

        If we think exclusively of footnotes, we will see no difference
        between what I have just called a {\em reference mark} and the {\em
        id}.  We clearly see the difference in other kinds of notes:
        Line notes, for example, have an id, but not a reference mark.

    \stopSmallPrint

  \item The {\em text} or {\em contents} of the note, always located at a
    different point on the page or in the document than the command that
    generates the note and indicates its content.

  \item The {\em label} associated with the note: a label or name
    associated with a note that is not shown in the final document, but
    allows us to refer to it and retrieve its ID elsewhere in the document.

\stopitemize

\startsubsection
    [title=Types of notes in \ConTeXt\ and commands associated with them]

We have various types of notes in \ConTeXt.  For the moment I will only
list them, describing them in general terms and providing information about
the commands that generate them.  Later I will develop the first two:

\startitemize

  \item {\bf Footnotes:} undoubtedly the most popular, to the extent that
    it is common for all types of notes to be referred to generically as
    {\em footnotes}.  Footnotes introduce a {\em mark} with the note's {\em
    id} at the point in the document where the command is found, and
    insert the text of the note itself at the bottom of the page where the
    mark appears.  They are created with the \tex{footnote} command.

  \item {\bf Endnotes:} these notes, created with the \tex{endnote}
    command, insert a marker with the note's identifier at the point in the
    document where the \tex{endnote} command is placed.  However, the
    note's contents are inserted at another point in the document, and the
    insertion is produced by a different command: (\tex{placenotes}).

  \item {\bf Margin notes:} as their name suggests, they are written in the
    margin of the text and there is no ID or automatically generated mark
    or anchor in the body of the document.  The two main commands (but not
    the only ones) that create them are \tex{inmargin} and \tex{margintext}.

  \item {\bf Line notes:} a type of note typical of environments where
    lines are numbered, such as in the case of
    \tex{startlinenumbering ... \stoplinenumbering}
    (see \in{Section}[sec:linenumbering]).  The note, which is usually
    written at the bottom, refers to a specific line number.  They are
    generated with the \PlaceMacro{linenote}\tex{linenote} command which is
    configured with \PlaceMacro{setuplinenote}\tex{setuplinenote}.  This
    command prints no {\em mark} in the body of the text, but in the note
    itself it prints the line number the note refers to (used as the {\em ID}).

\stopitemize

I will now exclusively develop the first two types of notes, because:

\startitemize

  \item margin notes are treated elsewhere (\in{Section}[sec:margintext]);
    and 

  \item line notes have a highly specialised use (especially in critical
    editions) and I believe that in an introductory document like this one,
    it is enough for the reader to know that they exist.

% JD note: this URL is dead.  I asked Pablo about it, and he told me that
% not only did some of the files (Flash videos!) disappear, but the
% material is now obsolete.  So, unless/until someone comes up with a
% suitable replacement, we will have to do without.
%  \startSmallPrint
%
%      However, for the interested reader I recommend a video (in Spanish)
%      accompanied by a text (also in Spanish) about critical editions in
%      \ConTeXt, the author of which is Pablo Rodríguez.  It is available at
%      \goto{this link}[url(http://www.ediciones-criticas.tk/)].  It is also
%      quite useful for understanding several of the general settings of
%      notes in general.
%
%  \stopSmallPrint

\stopitemize

\stopsubsection


\startsubsection
    [title=A close look at footnotes and endnotes]
\PlaceMacro{footnote}\PlaceMacro{endnote}

The syntax for the footnotes and endnotes commands and the configuration
and customisation mechanisms they have are quite similar, since, in
reality, both types of notes are particular instances of a more general
construction (notes), other instances of which can be created with the
\tex{definenote} command (see \in{Section}[sec:definenote]).

The syntax of the command that creates each of these kinds of notes is as
follows:

\starttyping
\footnote[Label]{Text}
\endnote[Label]{Text}
\stoptyping

where:

\startitemize

  \item {\tt Label} is an optional argument that assigns the note a label
    that will allow us to refer to it elsewhere in the document.

  \item {\tt Text} is the content of the note. It can be as long as we
    wish, and include special paragraphs and settings, although it should
    be noted that when it comes to footnotes, correct page layout is quite
    difficult in documents with abundant and excessively long notes.

    \startSmallPrint

% TODO: can someone list some examples of such things?  Or otherwise
% provide insight?

        In principle, any command that could be used in the main text can
        be used in the note text.  However, I have been able to verify that
        certain constructions and characters that do not pose any kind of
        problem in the main text, do generate a compilation error when they
        take place in the note text.  These cases I found as I was testing,
        but I have not organised them in any way.

    \stopSmallPrint

\stopitemize

When the {\tt Label} argument has been used to set a label for the note,
the \PlaceMacro{note}\tex{note} command allows us to retrieve the ID of the
note in question.  This command prints the ID of the note associated with
the label it takes as an argument on the document.  Thus, for
example:\footnote{Because footnotes are found at the bottom of the page,
    this example uses a rectangle around the typeset version on the right
    side to give an idea of how footnotes are arranged on a
    (landscape-orientation) page.}

\startbuffer[Humpty]
Humpty Dumpty\footnote[humpty]
{Probably the best-known English 
nursery rhyme character}
sat on a wall,\\
Humpty Dumpty\note[humpty]
had a great fall.\\
All the king's horses and
all the king's men\\
Couldn't put Humpty\note[humpty]
together again.
\stopbuffer

\ShowAndTypeset[Humpty]{height=130mm}

% TODO: this works in MkIV, but in LMTX the footnote goes to the bottom of
% the page.  Question in to the mailing list, hopefully answer soon.
% Same for example after this one.
%
% In any case, since the situation is no longer the same, this section
% needs a major re-write
%% 
%% \startDoubleExample
%% \startcolumns[n=2,distance=1.cm,separator=rule]
%% %\switchtobodyfont[small]
%% %\setupnotation[footnote][width=-1cm]
%% \starttyping
%% Humpty Dumpty\footnote[humpty]{Probably the
%% best-known English nursery rhyme character}
%% sat on a wall, Humpty Dumpty\note[humpty]
%% had a great fall.\\
%% All the king's horses and
%% all the king's men\\Couldn't put
%% Humpty\note[humpty] together again.
%% \stoptyping
%% \column
%% \switchtobodyfont[small]
%% \obeylines \setupwhitespace[none]
%% Humpty Dumpty\footnote[humpty]{Probably the best-known English nursery
%% rhyme character} sat on a wall,
%% Humpty Dumpty\note[humpty] had a great fall.
%% All the king's horses and all the king's men
%% Couldn't put Humpty\note[humpty] together again.
%% \stopcolumns
%% \stopDoubleExample

The main difference between \tex{footnote} and \tex{endnote} is the place
where the note appears:

\startdescription{\tex{footnote}}

As a rule, it prints the note text at the bottom of the page on which the
command is located, so that the note mark and its text (or the beginning of
the text, if it is to be spread over two pages) will appear on the same
page.  To do this, \ConTeXt\ will make the necessary adjustments in
typesetting the page by calculating the space required by the location of
the note at the bottom of the page.

\startSmallPrint

    Historical note: in \ConTeXt\ MkIV, in some
    environments, \tex{footnote} would insert the text of the note, not at
    the bottom of the page itself but at the \quotation{foot} of the
    environment.  This occurred, for example, in tables, and also in the
    {\tt columns} environment.  In such cases, if one wanted the notes
    inside the environment to be placed at the bottom of the page, instead
    of \tex{footnote} the \tex{footnotetext} command could be used in
    combination with the \tex{note} command mentioned above.  The former,
    which also accepts a label as an optional argument, prints only the
    note text but not the label.  But as \tex{note} prints only the label
    without the text, combining both allows us to place the note at the
    point wherever we want it.  So, for example, we could
    write \tex{note[MyLabel]} within a table or a multi-column environment,
    and then, once out of that
    environment, \type{\footnotetext[MyLabel]{Note text}}.

% TODO: this example doesn't make sense in LMTX.
%       But an example of where / when to use \footnotetext might be illustrative.
%       Maybe the opposite, showing a footnote at the bottom of a 2-column
%       environment.
%
%    Another example of the use of \tex{footnotetext} in combination
%    with \tex{note} would be notes inside other notes. For example:
%
%    \startDoubleExample
%    %\switchtobodyfont[small]
%    \setupnotation[footnote][width=-1cm]
%\vbox{%
%    \starttyping
%This%
%\footnote{or this\note[noteB], if you prefer.}%
%\footnotetext[noteB]
%{or possibly even this one\note[noteC].}
%\footnotetext[noteC]{could be something
%entirely different.}
%is a sentence with nested notes.
%    \stoptyping
%}
%
%    This%
%    \footnote{or this\note[noteB], if you prefer.}%
%    \footnotetext[noteB]{or possibly even this\note[noteC].}\footnotetext[noteC]{could be something
%        entirely different.}
%    is a sentence with nested notes.
%
%    \stopDoubleExample

\stopSmallPrint

\stopdescription

\startdescription{\tex{endnote}}

    only prints the note anchor at the point in the source file where the
    command is located.  The actual content of the note is inserted at
    another point in the document with another command,
    (\PlaceMacro{placenotes}\tex{placenotes[endnote]}) which, at the point
    where it is located, will insert the contents of {\em all} the endnotes
    of the document (or of the chapter or section in question).

\stopdescription

\stopsubsection


\startsubsection
  [
    reference=sec:localfootnotes,
    title={Local notes},
  ]
  \PlaceMacro{startlocalfootnotes}\PlaceMacro{placelocalfootnotes}

The \tex{startlocalfootnotes} environment means that the footnotes included
within it are considered to be {\em local} notes, which means that their
numbering will be reset and that the content of the notes will not be
automatically inserted along with the rest of the notes, but only at the
point in the document where the \tex{placelocalfootnotes} command is found,
which may or may not be within the environment.

% JD originally added this, but I'm not sure it adds a lot.  And it needs
% some formatting help (maybe make it into a \ShowAndTypeset?)
%% For example, here is an example of placing footnotes inside
%% the \quotation{double example} environment used in this introduction:
%% 
%% \startDoubleExample
%% \switchtobodyfont[small]
%% \setupnotation[footnote][width=-1cm] % This moves the footnote # a bit right
%% \vbox{\starttyping
%% \startlocalfootnotes
%% Humpty Dumpty\footnote[humpty]{Probably the
%% best-known English nursery rhyme character}
%% sat on a wall, Humpty Dumpty\note[humpty]
%% had a great fall.\\
%% All the king's horses and
%% all the king's men Couldn't put
%% Humpty\note[humpty] together again
%% \blank[3*line]\hrule
%% \placelocalfootnotes
%% \stoplocalfootnotes
%% \stoptyping}
%% 
%% \startlocalfootnotes
%% Humpty Dumpty\footnote[humpty]{Probably the best-known English nursery
%% rhyme character} sat on a wall,\\ Humpty Dumpty\note[humpty] had a great
%% fall.\\
%% All the king's horses and all the king's men\\
%% Couldn't put Humpty\note[humpty] together again
%% \blank[3*line]\hrule
%% {\placelocalfootnotes}
%% \stoplocalfootnotes
%% \stopDoubleExample
%% 
%% Note that \tex{placelocalfootnotes} doesn't format the local footnotes in
%% the same way footnotes are ordinarily formatted.  This example has explicit
%% command to skip some space (\type{\blank[3*line]}) and to draw a horizontal
%% line (\type{\hrule}). As you can see, \type{hrule} goes all the way across
%% the width of the text, unlike the short line used for normal footnotes.

\stopsubsection

\startsubsection
  [
    reference=sec:definenote,
    title={Creating and using customised types of notes},
  ]
  \PlaceMacro{definenote}

We can create special types of notes with the \tex{definenote} command.
This can be useful in complex documents where there are notes from
different authors, or for different purposes, to graphically distinguish
each of the types of notes in our document by means of a different format
and different numbering.

The syntax of \tex{definenote} is as follows:

\type{\definenote[Name][Model][Configuration]}

where:

\startitemize

  \item {\tt Name} is the name we assign to our new type of note.

  \item {\tt Model} is the note model that will be used initially. It can
    be {\tt footnote} or {\tt endnote}; in the former case our note model
    will work as footnotes, and in the latter case as endnotes, although to
    insert them in the document we would not use
    \PlaceMacro{placenotes}\tex{placenotes[endnote]}
    but \tex{placenotes[Name]} (the name we have assigned to these kinds of
    notes).

    \startSmallPrint

        In theory this argument is optional, although in my tests some
        notes created without it were not visible, and I have not had the
        patience to find out what the cause was.

    \stopSmallPrint

  \item {\tt Configuration} is an optional second argument that allows us
    to distinguish our new type of notes from its model: either by setting
    a different format, or a different type of numbering, or both.

    \startSmallPrint

        According to the official list of \ConTeXt\ commands (see
        \in{Section}[sec:qrc-setup-en]) the settings that can be provided
        when the new type of note is created are based on those that could
        be provided later with \tex{setupnote}.  However, as we shall see
        shortly, there are actually two possible commands for setting up
        notes: \tex{setupnote} and \tex{setupnotation}.  So I think it is
        preferable to omit this argument when creating the note type, and
        then set up our new notes using the appropriate commands. At least
        this is easier to explain.

    \stopSmallPrint

\stopitemize

For example, the following item will create a new note type called
\quotation{{\tt BlueNote}} that will be similar to footnotes but its contents
will be printed in bold and blue:

\starttyping
\definenote  [BlueNote] [footnote]
\setupnotation
  [BlueNote]
  [color=blue, style=bf]
\stoptyping

Once we have created a new note type, e.g., {\tt BlueNote}, the command
allowing us to use it will be available.  In our example this will be
\tex{BlueNote} the syntax of which will be similar to \tex{footnote}:

\type{\BlueNote[Label]{Text}}

\stopsubsection


% TODO: (see footnote below) count the number of options in LMTX.

\startsubsection
    [title=Configuring notes]
\PlaceMacro{setupnote}\PlaceMacro{setupnotation}

% TODO: the following footnote is extremely long.  Would it be better as
% \startSmallPrint text?

The configuration of notes (footnotes or endnotes, notes created with
\tex{definenote} and also line notes set up with \tex{linenote}) is
achieved with two commands: \tex{setupnote} and
\tex{setupnotation}\footnote{The \tex{setupnote} command has 35 {\em direct}
configuration options and 45 additional options inherited from
\tex{setupframed}; \tex{setupnotation} has 45 direct configuration options
and another 23 inherited from \PlaceMacro{setupcounter}\tex{setupcounter}.
Since these options are not documented and, although for many of them we
can work out their usefulness from their name, we need to check whether our
intuition is true or not.  Also, taking into account that many of these
options allow a number of values and they all have to be tested\unknown\
You will see that in order to write this explanation I had to do quite a
number of tests; and although doing a test is quick, doing a lot of tests
is slow and boring.  Therefore, I hope the reader will excuse me if I say
that in addition to the two general note configuration commands I mention
in the main text, and on which the following explanation focuses, there are
four other possible configuration commands that I will omit from
my explanation:

  \startitemize

    \item \PlaceMacro{setupnotes}\tex{setupnotes} and
      \PlaceMacro{setupnotations}\tex{setupnotations}: in other words, the
      same names as above, but in the plural. The wiki says that the singular and
      plural versions of the command are synonymous, and I believe it.

    \item \PlaceMacro{setupfootnotes}\tex{setupfootnotes} and
      \PlaceMacro{setupendnotes}\tex{setupendnotes}: these are supposed to
      be specific applications for footnotes and endnotes, respectively.
      Perhaps explaining the note setup on the basis of these commands
      would be easier; however, since the first option I tried
      with \tex{setupfootnotes} ({\tt numberconversion}) did not work,
      although I know that other options of these commands do work, I was
      too lazy to add the necessary tests (to the numerous tests I've
      already had to perform) to write what follows.
      \blank[small]
      What I do believe (from a few random tests) is that
      everything that works in these two commands, but whose explanation I
      am leaving out, also works in the commands for which I do give an
      explanation.

  \stopitemize
}.  The syntax for both is similar:

\starttyping
\setupnote[NoteType][Configuration]
\setupnotation[NoteType][Configuration]
\stoptyping

where {\tt NoteType} refers to the kind of note we are configuring ({\tt
footnote}, {\tt endnote} or the name of some note type we ourselves have
created), and {\tt Configuration} contains the particular configuration
options for the command.

The problem is that the names of these two commands do not make it very
clear what the difference is between them or what things each configures;
further, the fact that many of the options for these commands are not
documented does not help much either.  After a lot of testing I haven't
been able to reach any conclusion that would allow me to understand why
certain things are configured with one, while others are configured with
the other,\footnote{There is a page in the \goto{ \ConTeXt\ wiki}
    [url(https://wiki.contextgarden.net/Unexpected\_behavior)] that I
    discovered by chance (since it is not specifically dedicated to notes),
    which suggests that the difference is that \tex{setupnotation} controls
    the text of the note to be inserted, and \tex{setupnote} the
    environment of the note in which it will be placed~(?)  But this is
    inconsistent with the fact that, for example, the width of the note
    text (which has to do with its {\em insertion}) is controlled by the
    {\tt width} option of \tex{setupnote} and not by
    the \tex{setupnotation} option with the same name.  What is controlled
    here is the width of the space between the mark and the note text.}
except perhaps that, because of the choices I have made to make it
work, \tex{setupnotation} always affects the note text, or the ID that is
printed with the note text, while \tex{setupnote} has some options that
affect the mark for the note inserted in the main text.

I will now try to organise what I have found out after doing some tests
with the different options of both commands.  I leave most of the options
for both aside, as they are not documented and I have not been able to draw
any conclusions as to what they are for or under what conditions they
should be used:

\startitemize

  \starthead {\bf ID used for the mark:} notes are always identified by a
      number.  What we can configure here is:
  \stophead

  \startitemize

    \item {\em The first number}: controlled by {\tt start} in
      \tex{setupnotation}.  Its value has to be a whole number, and it uses
      this to begin counting notes.

    \item {\em The numbering system}: which depends on the {\tt
        numberconversion} option in \tex{setupnotation}.  Its values can be:

      \startitemize[packed]

        \item {\em Arabic numerals}: {\tt n}, {\tt N} or {\tt numbers}.

        \item {\em Roman numerals}: {\tt I}, {\tt R}, {\tt Romannumerals},
          {\tt i}, {\tt r}, {\tt romannumerals}.  The first three are upper
          case Roman numerals and the last three lower case.

        \item {\em Numbering with letters}: {\tt A}, {\tt Character}, {\tt
            Characters}, {\tt a}, {\tt character}, {\tt characters},
          depending on whether we want the letters to be in upper case (the
          first three options) or lower case (the rest).

        \item {\em Numbering with words}. In other words, we write the word
          that designates the number and so, for example, \quote{3} becomes
          \quotation{three}.  Two methods are possible; the {\tt Words}
          option writes the words in upper case and {\tt words} in lower
          case.

        \item {\em Numbering with symbols}: we can use four different sets
          of symbols depending on the option chosen: {\tt set~0}, {\tt
          set~1}, {\tt set~2} or {\tt set~3}.  On 
          \at{page}[examples of conversion set] there is an example of the
          symbols used in each of these options.

      \stopitemize

    \item {\em The event that determines restarting note numbering}: this
      depends on the {\tt way} option in \tex{setupnotation}. When the
      value is {\tt bytext} all notes in the document will be numbered
      sequentially without the numbering being reset. When it is {\tt
          bychapter}, {\tt bysection}, {\tt bysubsection}, etc., the note
      counter will be reset each time the chapter, section or subsection is
      changed, while when it is {\tt byblock} it resets the numbering each
      time we change blocks in the document macrostructure
      (see \in{Section}[sec:macrostructure]).  The {\tt bypage} value
      causes the note counter to restart each time the page is changed.

  \stopitemize

  \starthead {\bf Configuring the note mark:}
  \stophead

  \startitemize

    \item Whether or not to show it: {\tt number} option in \tex{setupnotation}.

    \item Placement of the mark in relation to the note text: the {\tt
        alternative} option in \tex{setupnotation}: it can take any of the
      following values: {\tt left}, {\tt inleft}, {\tt leftmargin}, {\tt
          right}, {\tt inright}, {\tt rightmargin}, {\tt inmargin}, {\tt
          margin}, {\tt innermargin}, {\tt outermargin}, {\tt serried},
      {\tt hanging}, {\tt top}, {\tt command}.

    \item Format of the mark in the note itself: the {\tt numbercommand}
      option in \tex{setupnotation}.

    \item Format of the mark in the body of the text: the {\tt textcommand}
      option in \tex{setupnote}.

      \startSmallPrint

          The {\tt numbercommand} and {\tt textcommand} options must
          consist of a command that takes the contents of the mark as an
          argument.  This can be a command we define ourselves.  However, I
          have found that simple formatting commands (\tex{bf}, \tex{it},
          etc.\null) work, even though they are not commands that take
          an argument.

      \stopSmallPrint

    \item Distance between the mark and the text (in the note itself): the
      {\tt distance} and {\tt width} options in \tex{setupnotation}.  I was
      unable to discover the difference (if indeed there is one) between
      using one or the other option.

    \item Existence or not of a hyperlink that allows jumping between the
      mark in the main text and the mark in the note itself: the {\tt
          interaction} option in \tex{setupnote}.  With {\tt yes} as a
      value there will be a link, and with {\tt no} there will not be one.

  \stopitemize

  \starthead {\bf Configuring the note text itself.} \stophead

  We can influence the following aspects:

  \startitemize

    \item Placement: this depends on the {\tt location} option
      in \tex{setupnote}.

      \startSmallPrint

          In principle we already know that footnotes are placed at the
          bottom of the page ({\tt location=page}) and endnotes at the
          point at which the \tex{placenotes[endnote]} ({\tt
          location=text}) command is found.  However, we can adjust
          this behavior and set, for example, {\tt location=text} which
          will cause footnotes to work similarly to endnotes, so they
          appear at the point in the document where the
          \tex{placenotes[footnote]} command is found (or the command
          specific to footnotes, \tex{placefootnotes}).  With this method
          we could, for example, print the notes under the paragraph in
          which they are found.

      \stopSmallPrint

    \item Paragraph separation between notes: by default each note is
      printed in its own paragraph, but we can have all the notes on the
      same page printed in the same paragraph by setting the {\tt
      paragraph} option in \tex{setupnote} to \MyKey{yes}.

    \item Style in which the note text itself will be written: the {\tt
        style} option in \tex{setupnotation}.

    \item Letter size: the {\tt bodyfont} option in \tex{setupnote}.

    \startSmallPrint

      This option is only for the case where we want to manually set a font
      size for the footnotes.  It is almost never a good idea to do this as,
      by default, \ConTeXt\ adjusts the font size of the footnotes so that
      it is smaller than the main text, but with a size {\em that is
      proportionate} to that of the font size in the main body.

    \stopSmallPrint

  \item Left margin for the note text: the {\tt margin} option in
    \tex{setupnotation}.

  \item Maximum width: the {\tt width} option in \tex{setupnote}.

  \item Number of columns: the {\tt n} option in \tex{setupnote} specifies
    that the note text will be in two or more columns.  The \quote{\tt n} value
    has to be a whole number.

  \stopitemize

  \item {\bf Space between notes or between notes and text:} here, we have
    the following options:

    \startitemize

      \item {\tt rule}, in \tex{setupnote} establishes whether or not there
        will be a line (rule) between the note area and the area of the
        page with the main text.  Its possible values are {\tt yes}, {\tt
        on}, {\tt no} and {\tt off}.  The first two enable the rule and
        the last two disable it.

      \item {\tt before}, in \tex{setupnotation}: command or commands to be
        run before inserting the note text.  Serves to insert additional
        spacing, dividing lines between notes, etc.

% TODO: I tried to see the effect of after with this command
% \setupnotation[before={\blank},after={\hpopo}] % or after={ZZZ} or after={\hrule}
% and while before works, I see no effect with after.  How does it work.

      \item {\tt after}, in \tex{setupnotation}: command or commands to be
        run after inserting the note text.

    \stopitemize

\stopitemize

\stopsubsection


\startsubsection
    [title={Temporary excluding notes when compiling}]
    \PlaceMacro{notesenabledfalse}\PlaceMacro{notesenabledtrue}

The \tex{notesenabledfalse} and \tex{notesenabledtrue} commands tell
\ConTeXt\ to enable or disable compiling of notes, respectively.  This
function can be useful if we wish to obtain a version without notes when
the document has numerous and extensive notes. In my personal experience,
for example, when I am correcting a doctoral thesis, I prefer to read it
the first time in one go, without the notes, and then do a second reading
with the notes incorporated.

\stopsubsection

\stopsection


\startsection
  [
    reference=sec:multiplecolumns,
    title={Paragraphs with multiple columns},
  ]

Typesetting the text in more than one column is a possibility that can be
established:

\startitemize[a]

  \item As a general feature of the page layout.

  \item As a feature of certain constructions such as, for example,
    structured lists, footnotes or endnotes.

  \item As a feature applied to particular paragraphs in a document.

\stopitemize

In any of these cases, most commands and environments will work perfectly
even if we are working with more than one column.  There are, however, some
limitations, mainly in relationship to floating objects in general
(see \in{Section}[sec:floating objects]) and with tables in particular
(\in{Section}[sec:tables]), even if they are not floats.

With regard to the number of columns allowed, \ConTeXt\ has no theoretical
limit.  However, there are two physical limits that have to be taken into
account:

\startitemize

  \item The width of the paper: an unlimited number of columns requires an
    unlimited width of paper (if the document is to be printed) or screen
    (if it is a document intended to be displayed on screen).  In practice,
    taking into account the {\em normal} width of the paper sizes that are
    marketed and used to make up books, and the screens of computer
    devices, it is difficult for a text made up of more than four or five
    columns to fit well.

  \item The size of the computer memory: the \ConTeXt\ reference manual
    points out that, in {\em normal} systems (neither particularly powerful
    nor particularly limited in resources), between 20 and 40 columns can
    be handled.

\stopitemize

In this section I will focus on the use of the multi-column mechanism in
special paragraphs or fragments, since:

\startitemize

  \item Multiple columns as a page layout option have already been
    discussed (in \in{Subsection}[sec:pages-columns] of
    \in{Section}[sec:pages-other-matters]).

  \item The possibility offered by certain constructions (such as
    structured lists or footnotes) to typeset text in more than one column
    is discussed in relation to the construct or environment in question.

\stopitemize

\stopcolumns


\startsubsection
  [title={The \tex{startcolumns} environment}]
  \PlaceMacro{startcolumns}

The normal procedure for inserting fragments made up of several columns
into a document is to use the {\tt columns} environment whose format is:

\type{\startcolumns[Configuration] ... \stopcolumns}

where {\tt Configuration} allows us to control many aspects of the
environment.  We can indicate the desired configuration each time we call
the environment, or adapt the default operation of the environment for all
calls to the environment; the latter is achieved with:

\PlaceMacro{setupcolumns}\type{\setupcolumns[Configuration]}

In both cases the configuration options are the same.  The most important
ones, ordered according to their function, are the following:

\startitemize

  \item {\bf Options that control the number of columns and the space
      between them:}

    \startitemize

      \item {\tt n}: controls the number of columns. If this is omitted,
        two columns will be generated.

      \item {\tt nleft}, {\tt nright}: these options are used in two-sided
        document layout (see \in{Subsection}[sec:double-sided] of
        \in{Section}[sec:pages-other-matters]), to establish the number of
        columns on left (even) and right (odd) pages respectively.

      \item {\tt distance}: specifies the space between columns.

      \item {\tt separator}: determines what marks the separation between
        columns.  It can be {\tt space} (default value) or {\tt rule} in
        which case a line (rule) will be generated between the columns. In
        the event that a rule is established between columns, this rule can
        in turn be configured with the following two options:

        \startitemize

          \item {\tt rulecolor}: colour of the line.

          \item {\tt rulethickness}: thickness of the line.

        \stopitemize

      \item {\tt maxwidth}: specifies the maximum width that the columns
        and the space between them may occupy.

    \stopitemize

  \item {\bf Options that control text distribution in columns:}

    \startitemize

      \item {\tt balance}: by default, \ConTeXt\ {\em balances} columns,
        meaning it distributes the text between them so that they have more
        or less the same amount of text.  However, we can set this option
        with the \quotation{{\tt no}} the text will not start in a column
        until the previous one is full.

      \item {\tt direction}: determines in which direction the text is
        distributed between the columns. By default, the natural reading
        order is followed (from left to right), but giving this option the
        {\tt reverse} value results in right to left.

    \stopitemize

    \starthead {\bf Options affecting typesetting of text within this
environment:}
    \stophead

    \startitemize

      \item {\tt tolerance}: text written in more than one column means
        that line width within a column is smaller, and as explained when
        describing the mechanism \ConTeXt\ uses for constructing lines
        (\in{Section}[sec:lines]), this makes it difficult to locate
        optimal points for inserting line breaks.  This option allows us to
        temporarily alter the horizontal tolerance in an environment (see
        \in{Section}[sec:horizontaltolerance]), to facilitate the
        typesetting of the text.

      \item {\tt align}: this controls the horizontal alignment of lines
        within the environment.  It can take any of the following values:
        {\tt right}, {\tt flushright}, {\tt left}, {\tt flushleft}, {\tt
            inner}, {\tt flushinner}, {\tt outer}, {\tt flushouter}, {\tt
            middle} or {\tt broad}.  For the meaning of these options,
        see \in{Section}[sec:setupalign].

      \item {\tt color}: specifies the name of the colour in which the text
        within the environment will be written.

  \stopitemize

\stopitemize

Another way to affect the distribution of text among columns is
the \tex{column} command.  Using this command at any place in the body of
the text instructs \ConTeXt\ to stop the current column and to continue in
the next column.

For example, the following demonstrates how one could typeset some Spanish
text (from the original Spanish version of ANSS) next to a translation into
English.  The English part was intentionally truncated to demonstrate how
the \tex{column} command prevents the normal balancing of the columns.  (In
this case the full English translation would occupy the same number of
lines as the Spanish.)

\startbuffer
\startnarrower
  \startcolumns[distance=1cm,separator=rule]
      La anchura del papel: Un número ilimitado de columnas exige una anchura
      ilimitada de papel (si el documento se va a imprimir) o de pantalla (si es
      un documento pensado para ser mostrado en pantalla).
   \column 
      The width of the paper: An unlimited number of columns requires a
      unlimited width of paper \unknown
  \stopcolumns
\stopnarrower
\stopbuffer

\need 5\baselineskip
\typebuffer
{\switchtobodyfont[small]
\getbuffer
}

\stopsubsection

\startsubsection
    [title={Parallel paragraphs}]
    \PlaceMacro{defineparagraphs}\PlaceMacro{setupparagraphs}

A specific version of the multi-column composition is parallel paragraphs.
In this type of construction the text is distributed across two columns
(usually, although sometimes more than two), but it is not allowed to flow
freely between them, and instead maintains strict control over what will
appear in each column.  This is very useful, for example, to generate
documents which contrast two versions of a text, such as the new and the
old version of a recently amended law, or in bilingual editions; or also to
write glossaries for specific text definitions where the text to be defined
appears on the left and the definition on the right, etc.

Normally a table mechanism is used to process these kinds of paragraphs,
but this is because most text processors are not as powerful as \ConTeXt,
which has the \tex{defineparagraphs} and \tex{setupparagraphs} commands.
These commands allow the construction of parallel paragraphs using the
column mechanism, which, although it has limitations, is more flexible than
the table mechanism.

\startSmallPrint

    As far as I know these paragraphs have no special name.  I have called
    them \quotation{parallel paragraphs} because to me it seems to be a
    more descriptive term than the one \ConTeXt\ uses to refer to them:
    \quotation{{\em paragraphs}}.

\stopSmallPrint

The basic command here is \tex{defineparagraphs} whose syntax is:

\type{\defineparagraphs[Name][Configuration]}

where {\tt Name} is the name we give this construction, and {\tt
Configuration} specifies the features it will have, which can also be
set later with

\type{\setupparagraphs[Name][Column][Configuration]}

where {\tt Name} is the name given when creating it, {\tt Column} is an
optional argument allowing us to configure a particular column, and {\tt
Configuration} allows us to determine how it works in practice.

For example:

\starttyping
\defineparagraphs
  [MurciaFacts]
  [n=3, before={\blank},after={\blank}]

\setupparagraphs
  [MurciaFacts][1]
  [width=.1\textwidth, style=bold]

\setupparagraphs
  [MurciaFacts][2]
  [width=.4\textwidth]
\stoptyping

The above fragment would create a three-column environment called
MurciaFacts and then set the first column to take up 10 percent of the line
width and be written in bold, and set the second column to take up 40
percent of the line width. As the third column is not configured, it will
have the remaining width, i.e., 50\%.

Once the environment has been created, we can use it to write a brief
history of Murcia:

\vbox{\starttyping
  \startMurciaFacts
    825
  \MurciaFacts
    City of Murcia founded.
  \MurciaFacts
    The origins of the city of Murcia are uncertain, but there is evidence
    that it was ordered to be founded under the name of Madina (or Medina)
    Mursiya in the year 825 by the Emir of al-Àndalus Abderramán II,
    probably built over a much earlier settlement.
  \stopMurciaFacts
\stoptyping}

\defineparagraphs
  [MurciaFacts]
  [n=3, before={\blank},after={\blank}]

\setupparagraphs
  [MurciaFacts][1]
  [width=.1\textwidth, style=bold]

\setupparagraphs
  [MurciaFacts][2]
  [width=.4\textwidth]

\example{\startMurciaFacts
    825
  \MurciaFacts
    City of Murcia founded.
  \MurciaFacts
    The origins of the city of Murcia are uncertain, but there is evidence
    that it was ordered to be founded under the name of Madina (or Medina)
    Mursiya in the year 825 by the Emir of al-Àndalus Abderramán II,
    probably built over a much earlier settlement.
  \stopMurciaFacts}

If we wanted to continue telling the story of Murcia, a new instance of the
environment (\tex{startMurciaFacts}) would be needed for the next event,
because it is not possible to include several {\em rows} with this
mechanism.  (That is, this particular environment doesn't allow us to
continue with another row telling what happened in, for example, the year
830.)

From the example just given, I would like to highlight the following details:

\startitemize

  \item Once an environment has been created with, say, the command 
    \tex{defineparagraphs[MaryPoppins]}, this becomes a normal environment
    which starts with \tex{startMaryPoppins} and ends with
    \tex{stopMaryPoppins}.

  \item A \tex{MaryPoppins} command is also created, used within the
    environment to indicate when to change the column.

\stopitemize

\startSmallPrint 

    The example above of the \quotation{plain} \tex{startcolumns}
    environment could have been an example for this \quotation{parallel
        paragraphs} section.  However, it seemed like the correct place to
    introduce the \tex{column} command.  The \ConTeXt\ wiki notes
    that \tex{column} is an alias for the \tex{columnbreak} command, which
    may even be an easier to remember name.  Note that the command is
    not \tex{columns}, which is a slight inconsistency, compared to the
    analogous command created by \tex{defineparagraphs[...]}.

\stopSmallPrint

As for the configuration options for parallel paragraphs
(\tex{setupparagraphs}), I understand that, at this stage of the
introduction, and taking into account the example just given, the reader is
already prepared to work out the purpose of each of the options.
Therefore, below I will only indicate the name and type of the options and,
where appropriate, the possible values.  Remember, though, that
\tex{setupparagraphs [Name] [Configuration]} sets up configurations that
affect the whole environment, while \tex{setupparagraphs [Name] [NumColumn]
[Configuration]} applies configurations exclusively to the column
indicated.

\startitemize[columns, three, packed]\switchtobodyfont[small]

  \item {\tt n}: Number

  \item {\tt before}: Command

  \item {\tt after}: Command

  \item {\tt width}: Dimension

  \item {\tt distance}: Dimension

  \item {\tt align}: Derived from \\\tex{setupalign}

  \item {\tt inner}: Command

  \item {\tt rule}: {\tt on} or  {\tt off}

  \item {\tt rulecolor}: Colour

  \item {\tt rulethickness}: Dimension

  \item {\tt style}: Style Command

  \item {\tt color}: Colour

\stopitemize

\startSmallPrint

    The above list of options is not complete; I have excluded from the
    list of options those that I would not normally explain here.  I have
    also taken advantage of the fact that we are in the section dedicated
    to columns to show the list of options in triple columns, although I
    have not done it with any of the commands explained in this section;
    instead, the {\tt columns} option of the {\tt itemize} environment was
    used, to which the next section is dedicated.

\stopSmallPrint

\stopsubsection

\stopsection


\startsection
  [
    reference=sec:itemize,
    title={Structured lists},
  ]

When information is presented in an orderly manner, it is easier for the
reader to grasp.  But if the arrangement is also visually perceptible, then
it highlights for the reader the fact that here we have a structured text.
This is why there are certain {\em constructions} or {\em mechanisms} that
try to visually organize the text, thus contributing to its structuring.
Of the tools that \ConTeXt\ makes available to the author for this purpose,
the most important one, which is the subject of this section, is the {\tt
itemize} environment that is used to develop what we could call {\em
structured lists}.

Lists consist of a sequence of {\em text elements} (which I will call
{\em items}), in which each item is preceded by a character that helps to
highlight it by differentiating it from the rest, and which I will call the
\quotation{separator}.  The separator can be a number, letter or symbol.
Often (but not always) the {\em items} are paragraphs, and the list is
formatted to ensure the {\em visibility} of the separator for each element;
usually by applying a hanging indent that makes it stand out\footnote{In
    typography, an indent that applies to all the lines of a paragraph
    except the first one is called a {\em hanging indent}, which makes the
    first word or character of the paragraph easy to find.}.  In the case
of nested lists, the indentation of each sub-list is gradually increased.
In HTML, lists where the separator is a number or character that increases
sequentially are called {\em ordered lists}, meaning that each {\em item}
of the list will have a different separator that will allow us to refer to
each element by its number or identifier; and it gives the name {\em
    unordered lists} to lists where the same character or symbol is used
for every item in the list.

\ConTeXt\ automatically manages the numbering or alphabetical sequencing of
the separator in numbered lists, as well as the indentation that nested
lists need to have.  Further, in the case of nesting unordered
lists, \ConTeXt\ also looks after the selection of a different character or
symbol that allows the level of a list {\em item} to be distinguished at a
glance, according to the symbol that precedes it.

\startSmallPrint

    The reference manual says that the maximum level of nesting in lists is
    four, but I guess that was the case in 2013, when the manual was
    written.  According to my tests there seems to be no limit to the
    nesting of {\em ordered} lists (in my tests I reached up to 15 levels
    of nesting).  While for unordered lists, there does not seem to be a
    limit either, in the sense that no matter how many nests we include, no
    error will be generated; but, for unordered lists, \ConTeXt\ only
    applies default symbols for the first nine levels of nesting.

    In any case, it should be pointed out that the excessive use of nesting
    in lists can have the opposite effect to what we intend, and that is
    that the reader feels lost, unable to locate each item in the general
    structure of the list.  For this reason I personally believe that while
    lists are a powerful tool for structuring a text, it is almost never a
    good idea to go beyond the third level of nesting; and even the third
    level should only be used in certain cases where we can justify it.

\stopSmallPrint

The general tool for writing lists in \ConTeXt\ is the \tex{itemize}
environment whose syntax is as follows:

\PlaceMacro{startitemize}\type{\startitemize[Options][Configuration] ... \stopitemize}

where the two arguments are optional.  The first one allows symbolic names
as content that have been assigned a precise meaning by \ConTeXt; the
second argument, which is rarely used, makes it possible to assign specific
values to certain variables that affect the functioning of the environment.

\startsubsection
  [
    reference=sec:itemize_select-list-type,
    title={Selection the kind of list and separator between {\em items}},
  ]

\startsubsubsection
  [title={Unordered lists}]

By default, the list generated by {\tt itemize} is an unordered list, in
which the separator will be automatically selected depending on the nesting
level:

\need 4\baselineskip

\startitemize[packed, columns, two]\switchtobodyfont[small]

  \sym{\convertnumber{set 0}{1}} For the first level of nesting.

  \sym{\convertnumber{set 0}{2}} For the second level of nesting.

  \sym{\convertnumber{set 0}{3}} For the third level of nesting.

  \sym{\convertnumber{set 0}{4}} For the fourth level of nesting.

  \sym{\convertnumber{set 0}{5}} For the fifth level of nesting.

  \sym{\convertnumber{set 0}{6}} For the sixth level of nesting.

  \sym{\convertnumber{set 0}{7}} For the seventh level of nesting.

  \sym{\convertnumber{set 0}{8}} For the eighth level of nesting.

  \sym{\convertnumber{set 0}{9}} For the ninth level of nesting.

\stopitemize

However, we can expressly indicate that we want the symbol associated with
a particular level to be used, simply by passing on the level number as an
argument. Thus, \tex{startitemize[4]} will generate an unordered list in
which the \triangleright\, character will be used as a separator,
regardless of the nesting level of the list.

We can also modify the predetermined symbol for each level
with \PlaceMacro{definesymbol}\tex{definesymbol}:

\type{\definesymbol[Level]{Symbol associated with the level}}

For example

\type{\definesymbol[1][\diamond]}

will cause the first level of unordered lists to use a \diamond\,
symbol. With this same command we can assign some symbols to nesting levels
higher than nine. Thus, for example

\type{\definesymbol[10][\copyright]}

will assign the international {\em copyright} symbol: \copyright\, to
nesting level 10.

\stopsubsubsection

\startsubsubsection
  [title=Ordered lists]

To generate an ordered list we need to tell \tex{startitemize} the kind of
ordering we want.  It can be:

\startitemize[intro, packed, 2*broad, columns, three]
\switchtobodyfont[small]

  \sym{{\bf n}} 1, 2, 3, 4, \unknown

  \sym{{\bf m}} {\os 1, 2, 3, 4, \unknown}

  \sym{{\bf g}} \alpha, \beta, \gamma, \delta, \unknown

  \sym{{\bf G}} \Alpha, \Beta, \Gamma, \Delta, \unknown

  \sym{{\bf a}} a, b, c, d, \unknown

  \sym{{\bf A}} A, B, C, D, \unknown

  \sym{{\bf KA}} \cap{a, b, c, d, \unknown}

  \sym{}

  \sym{{\bf r}} i, ii, iii, iv, \unknown

  \sym{{\bf R}} I, II, III, IV, \unknown

  \sym{{\bf KR}} \cap{i, ii, iii, iv, \unknown}

\stopitemize

The difference between {\tt n} and {\tt m} lies in the style of digits used
to represent the number: {\tt n} uses digits from the current font, while
the digits that {\tt m} uses are obtained with the \tex{os}
(\quotation{old-style}) command; \type{{\os 0 1 2 3 4 5 6 7 8 9}} produces
{\os 0 1 2 3 4 5 6 7 8 9}.

\stopsubsubsection

\stopsubsection


\startsubsection
  [
    reference=sec:itemize_item-type,
    title={Entering the items of a list},
  ]

As a rule, the items in a list created with \tex{startitemize} are
introduced with the \PlaceMacro{item}\tex{item} command;
however, \tex{item} also has an \quotation{environment version} form:
\PlaceMacro{startitem}\tex{startitem ... \stopitem}.  Thus the following example:

\startbuffer
\startitemize[a, packed]
\startitem First element \stopitem
\startitem Second element \stopitem
\startitem Third element \stopitem
\stopitemize
\stopbuffer
\ShowAndExecute

produces exactly the same result as

\startbuffer
\startitemize[a, packed]
\item First element
\item Second element
\item Third element
\stopitemize
\stopbuffer
\ShowAndExecute

Either \tex{item} or \tex{startitem} is the {\em general} command for
introducing an item into the list.  Along with those, there are the
following additional commands for when we want to achieve a special result:

\startitemize[3*broad]

  \sym{\PlaceMacro{head}\tex{head}} This command should be used in place of
    \tex{item} when we want to avoid inserting a page break after the item
    in question.

  \startSmallPrint

      A common construction is to include a nested list or a text block
      immediately below a list element, so that the list element, in a
      sense, functions as the {\em title} of the sub-list or text block.  In
      these cases a page break between that element and the subsequent
      paragraphs would be inadvisable.  If we use \tex{head} instead
      of \tex{item} to begin these elements \ConTeXt\ {\em will endeavour}
      (as far as possible) not to separate such an element from the next
      block.

  \stopSmallPrint

  \sym{\PlaceMacro{sym}\tex{sym}} The \type{\sym{Text}} command begins an
    item in which the argument text of \tex{sym} is used as the {\em
        separator}, rather than the usual number or symbol.  This list, for
    example, is constructed with items begin with \tex{sym} instead
    of \tex{item}.

  \sym{\PlaceMacro{sub}\tex{sub}} This command, which should be used only
    in ordered lists (where each item is preceded by a number or letter in
    alphabetical sequence), causes its item to keep the number of the
    previous item, and in order to indicate that the number is repeated and
    that this is not a mistake, the \quote{+} sign is printed on the
    left. This can be useful if we are referring to a previous list for
    which we suggest modifications but where, for the sake of clarity, the
    numbering of the original list should be maintained.

  \sym{\PlaceMacro{mar}\tex{mar}} This command maintains the numbering of
    the items, but adds a letter or character in the margin (which is
    passed to it as an argument, between curly brackets). I'm not quite
    sure how useful it is.

\stopitemize

There are two additional commands for introducing items, whose combination
produces very {\em interesting} effects and, if I may say so, I think it is
better to explain them with an example. These are \PlaceMacro{ran}\tex{ran}
(abbreviation of {\em range}) and \PlaceMacro{its}\tex{its}, abbreviation
of {\em items}.  The first one takes an argument (between curly brackets)
and the second repeats the symbol used as a separator in the list x number
of times (by default four times, but we can alter that by
using \tex{startitemize}'s {\tt items=} option).  The following example
shows how these two commands can work together to create a list that mimics
a questionnaire.  Note how {\tt width} is used to specify the width of
the \quotation{No \qquad Yes} column and {\tt distance} is used to specify
the gap between the check-boxes and the questions.  This is explained in
more detail in \in{Section}[sec:itemize_arg2].

\vbox{\starttyping
After reading the following introduction, answer the following questions:

\startitemize[8, packed][width=8em, distance=2em, items=5]
\ran{No \hss Yes}
\its I will never use \ConTeXt, it is too difficult.
\its I will only use it for writing big books.
\its I will always use it.
\its I like it so much I will call my next child \quotation{Hans}, as a tribute
to Hans Hagen.
\stopitemize
\stoptyping}


\need5\baselineskip
%\startlocalfootnotes
\example{After reading this introduction, answer the following questions:

\startitemize[8, packed][width=8em, distance=2em, items=5]
  \ran{No \hss Yes}
  \its I will never use \ConTeXt, it is too difficult.
  \its I will only use it for writing big books.
  \its I will always use it.
  \its I like it so much I will call my next child \quotation{Hans}, as a
    tribute to Hans Hagen.
\stopitemize
%\placelocalfootnotes
}
%\stoplocalfootnotes

It is possible to change the number of separators for an individual item by
giving \tex{its} an argument, as seen in the following example.

\startbuffer
    \startitemize[5,packed][width=8em, distance=1.5em, items=4]
      \ran{A bit\hss A lot}
      \its I like cake.
      \its I like pie.
      \its[3] I like cookies.
    \stopitemize
\stopbuffer

\typebuffer
\example{\getbuffer}

\stopsubsection


\startsubsection
  [
    reference=sec:itemize_arg1,
    title={Basic list configuration},
  ]

Recall that \tex{startitemize} allows for two arguments.  We have already
seen how the first argument lets us select the type of list we want.  But
we can also use it to indicate other characteristics of the list; this is
done through the following options for \tex{startitemize} in its first
argument:

\startitemize

  \item {\tt columns}: this option determines that the list is made up of
    two or more columns. After the columns option, the desired number of
    columns must be written as words separated by a comma: two, three,
    four, five, six, seven, eight or nine. If {\tt columns} is not followed
    by any number two columns are generated.

% TODO: is this (serried) really true?  Didn't work for me right here.

  \item {\tt intro}: this option tries not to separate the list, by a line
    break, from the paragraph that precedes it.

  \item {\tt continue}: in ordered lists (numerical or alphabetical) this
    option causes the list to continue the numbering from the last numbered
    list.  If the {\tt continue} option is used, it is not necessary to
    indicate what type of list we want, as it is assumed that it will be
    the same as the last numbered list.

  \item {\tt packed}: is one of the most used options.  Its use causes the
    vertical space between the different {\em items} on the list to be
    reduced as far as possible.

  \item {\tt nowhite:} produces an effect similar to {\tt packed}, but more
    drastic: not only does it reduce the vertical space between the items,
    but also the vertical space between the list and the surrounding text.

  \item {\tt broad}: increases the horizontal space between the item
    separator and the item text.  The space can be increased by multiplying
    a number by {\tt broad} as in, for example \type{\startitemize[2*broad]}.

% TODO: is this (serried) really true?  Didn't work for me right here.

  \item {\tt serried}: removes the horizontal space between the item
    separator and the text.

  \item {\tt intext}: removes the hanging indent.

  \item {\tt text}: removes the hanging indent and reduces vertical space
    between items.

  \item {\tt repeat}: in nested lists makes the numbering of a child level
    {\em repeat} the same level as the previous level.  This way we would
    have, on the first level: 1, 2, 3, 4; on the second level: 1.1, 1.2,
    1.3, etc.  The option must be indicated for the inside list, not on the
    outer list.

  \item {\tt margin, inmargin}: by default the list separator is printed on
    the left, but within the text area itself ({\tt\quotation{atmargin}}).
    The options {\tt margin} and {\tt inmargin} move the separator to the
    margin.

\stopitemize

\stopsubsection


\startsubsection
  [
    reference=sec:itemize_arg2,
    title={Additional list configuration},
  ]

The second argument, also optional, in \tex{startitemize} allows for a more
detailed and thorough configuration of lists.

\startitemize

  \item {\tt before, after}: commands to be run before starting or after
    closing the itemize environment, respectively.

  \item {\tt inbetween}: a command to be run between two {\tt items}.

  \item {\tt beforehead, afterhead}: a command to be run before or after an
    item input with the \tex{head} command.

  \item {\tt left, right}: a character to be printed to the left or right of
    the separator.  For example, to get alphabetical lists in which the
    letters are surrounded by parentheses we would write:

    \type{\startitemize[a][left=(, right=)]}

  \item {\tt stopper}: indicates a character to be written after the
    separator.  Only works in ordered lists.

  \item {\tt width, maxwidth}: width of the space reserved for the
    separator and, therefore, for the hanging indent.

  \item {\tt factor}: representative number of the separation factor
    between the separator and the text.

  \item {\tt distance}: measure of the distance between the separator and
    the text.

  \item {\tt leftmargin, rightmargin, margin}: margin to be added to the
    left (leftmargin) or right (rightmargin) of the items.

  \item {\tt start}: number from which the numbering of items will start.

  \item {\tt symalign, itemalign, align}: alignment of items.  Allows for
    the same values as \tex{setupalign}; {\tt symalign} controls alignment
    of the separator, {\tt itemalign} the item text, and {\tt align}
    alignment of both.

  \item {\tt identing}: indentation of the first line in the paragraphs
    within the environment.  See \in{Section}[sec:indentation]

  \item {\tt indentnext}: indicates whether the paragraph after the
    environment should be indented or not.  Values are {\tt yes}, {\tt no} and
    {\tt auto}.

  \item {\tt items}: in items entered input with \tex{its}, indicates the
    number of times the separator must be reproduced.

  \item {\tt style}, {\tt color}; {\tt headstyle}, {\tt headcolor}; {\tt
      marstyle}, {\tt marcolor}; {\tt symstyle}, {\tt symcolor}: these
    pairs of options control the style and colour of the items, according
    to whether they are introduced with \tex{item}, \tex{head}, \tex{mar}
    or \tex{sym} commands, respectively.

\stopitemize

\stopsubsection


\startsubsection
  [
    reference=sec:items command,
    title={Simple lists with the \tex{items} command},
  ]
  \PlaceMacro{items}

An alternative to the {\tt itemize} environment for very simple unnumbered
lists, where the items are not too big, is the \tex{items} command whose
syntax is:

\type{\items[Configuration]{Item 1, Item 2, ..., item n}}

The different list items are separated from each other by commas.  For
example:

\startbuffer
Graphics files can
have, among other things, the
following extensions:

\items{png, jpg, tiff, gif}
\stopbuffer
\ShowAndExecute

The configuration options supported by this command are a subset of the
{\tt itemize} ones, except for two specific options for this command:

\startitemize

  \item {\tt symbol}: this option determines the type of list that will be
    generated.  It supports the same values used for {\tt itemize} to select
    some type of list.

  \item {\tt n}: this option indicates from which item number there will be
    a separator.

\stopitemize

\stopsubsection


\startsubsection
  [
    reference=sec:setupitemize,
    title={Predetermining list behaviour and creating our own list types},
  ]

In the previous sections we have seen how to indicate what type of list we
want and what characteristics it should have.  But doing that every time a
list is called is inefficient and will usually produce an incoherent
document in which each list has its own appearance, but without the
different appearances meeting any criteria.

Preferable result for this:

\startitemize

  \item Predetermine the {\em normal} behaviour of the {\tt itemize}
    environment and \tex{items} in the document's preamble.

  \item Create our own customised lists.  For example: an alphabetically
    numbered list (which we could call {\tt ListAlpha}), a list numbered
    with Roman numerals ({\tt ListRoman}), etc.

\stopitemize

We achieve the first with the \tex{setupitemize} and \tex{setupitems}
commands.  The second requires the use of either the
\PlaceMacro{defineitemgroup}\tex{defineitemgroup} or the
\PlaceMacro{defineitems}\tex{defineitems} command.  The first will create a
list environment similar to {\tt itemize} and the second a command similar
to \tex{items}.

\stopsubsection

\stopsection


\startsection
    [title={Descriptions and enumerations}]

Descriptions and enumerations are two constructions that allow for the
consistent typesetting of paragraphs or groups of paragraphs that develop,
describe, or define a phrase or word.

\startsubsection
    [title={Descriptions}]

For descriptions we differentiate between a {\em title} and its explanation
or development.  We can create a new description with:

\PlaceMacro{definedescription}\type{\definedescription [Name] [Configuration]}

where {\tt Name} is the name by which this new construction will be known,
and {\tt Configuration} controls what our new structure will look like. After the
previous statement we will have a new command and an environment with the
name we have chosen. Thus:

\type{\definedescription [Concept]}

will create the commands:

\starttyping
  \Concept{Title}
  \startConcept {Title}  \stopConcept
\stoptyping

We will use the command for the case where the explanatory text of the
title consists of only one paragraph, and the environment for titles whose
description occupies more than one paragraph.  When the command is used,
the paragraph immediately following it is the one that will be considered
the title's explanatory text.  But when the environment is used, all
content will be formatted with the appropriate indentation to make it clear
how it relates to the title.

For example:

\starttyping
\definedescription
  [Concept]
  [alternative=left, width=1cm, headstyle=bold]

\Concept{Contextualise}

Place something in a certain context, or typeset a text with the typesetting
system called \ConTeXt.  The ability to correctly contextualise in any situation
is considered a sign of intelligence and good sense.

\stoptyping

This will generate the following result:

\startcolor[red]
    
\definedescription
  [Concept]
  [alternative=left, width=3cm, headstyle=bold]

\Concept{Contextualise}

Place something in a certain context, or typeset a text with the typesetting
system called \ConTeXt.  The ability to correctly contextualise in any situation 
is considered a sign of intelligence and good sense.

\stopcolor

As is normally the case with \ConTeXt, the appearance that our new
construction will have can be indicated at the time of its creation, with
the {\tt Configuration} argument or later on with \tex{setupdescription}:

\PlaceMacro{setupdescription}\type{\setupdescription [Name] [Configuration]}

where {\tt Name} is the name of our new description, and {\tt
Configuration} determines what it looks like.  Among the different
possible configuration options I will highlight:

\startitemize

  \item {\tt alternative}: This option is the one that fundamentally
    controls the appearance of the construction.  It determines the
    placement of the title in relation to its description. Its possible
    values are {\tt left}, {\tt right}, {\tt inmargin}, {\tt inleft}, {\tt
        inright}, {\tt margin}, {\tt leftmargin}, {\tt rightmargin}, {\tt
        innermargin}, {\tt outermargin}, {\tt serried} and {\tt hanging}.
    Their names are clear enough to get an idea of the result, although, in
    case of doubt, it is best to do a test to see how it looks.

  \item {\tt width}: controls the width of the box in which the title will
    sbe written.  Depending on the value of {\tt alternative} that distance
    will also be part of the indentation with which the explanatory text is
    written.

  \item {\tt distance}: controls the distance between the title and the
    explanation.

  \item {\tt headstyle}, {\tt headcolor}, {\tt headcommand}: these affect
    how the title itself will look.  With {\tt headcommand} we can indicate
    a command to which the title text will be passed as an argument.  For
    example: {\tt headcommand=\ttbs WORD} will make sure that the title
    text is in upper case.

  \item {\tt style}, {\tt color}: controls the appearance of the 
    descriptive text.

\stopitemize

\stopsubsection


\startsubsection
    [title={Enumerations}]

Enumerations are numbered text elements structured on several levels.  Each
element starts with a title that consists, by default, of the name of the
structure and its number, although we can change the title with the {\tt
text} option.  They are quite similar to descriptions, although they differ
in that:

\startitemize

  \item All the elements in an enumeration share the same title.

  \item Therefore they differ from each other by their numbering.

\stopitemize

This structure can be very useful, for example, to write formulas, problems
or exercises in a textbook, ensuring that they are numbered correctly and
formatted in a consistent manner.

We create an enumeration with

\PlaceMacro{defineenumeration}\type{\defineenumeration [Name] [Configuration]}

where {\em Name} is the name of the new construction, and {\em
Configuration} controls what it will look like.

So, in the following example:

\starttyping
  \defineenumeration
    [Exercise]
    [alternative=top, before=\blank, after=\blank, between=\blank]
\stoptyping

We have created a new structure called {\em Exercise} and, as happens with
enumerations, we will have the following new commands available:

\starttyping
  \Exercise
  \startExercise  
  \stopExercise
\stoptyping

The command is used just for a single paragraph {\em exercise}, and the
environment for multiple paragraph {\em exercises}. But since enumerations
can be up to four levels deep, the following commands and environments will
also be created:

\starttyping
  \subExercise
  \startsubExercise
  \stopsubExercise
Y, para controlar la numeración, los siguientes comandos adicionales  \subsubExercise
  \startsubsubExercise
  \stopsubsubExercise
  \subsubsubExercise
  \startsubsubsubExercise
  \stopsubsubsubExercise
\stoptyping

And, to control the numbering, the following additional commands:

% TODO: would the items below be clearer if they were setExercise,
% resetExercise, and nextExercise?

\startitemize

  \item \tex{setEnumerationName}: sets the current numbering value.

  \item \tex{resetEnumerationame}: sets the enumeration counter to zero.

  \item \tex{nextEnumerationName}: increases the enumeration counter by one.

\stopitemize

The appearance of enumerations can be determined at the time of their
creation or later with \tex{setupenumeration} whose format is:

\PlaceMacro{setupenumeration}\type{\setupenumeration [Name] [Configuration]}.

For each enumeration we can configure each of its levels separately. Thus,
for example, \tex{setupenumeration [subExercise] [Configuration]} will
affect the second level of the enumeration called \quotation{Exercise}.

The options and values configurable with \tex{setupenumeration} are similar
to those in \tex{setupdescription}.

% TODO: show an example of setting up an exercise and some of the 

\stopsubsection

\stopsection

\startsection
  [
    reference=sec:FramesLines,
    title={Lines and frames},
  ]

% TODO I'm not sure what exactly the ConTeXt manual said, but is JAL
% confusing what the manual said about "plain TeX's capabilities" vs.
% ConTeXt's capabilities?  If so, the beginning of this paragraph needs
% some rewording.
% And "graphic" may be confusing me, because I'm not sure if that word
% makes people think of lines and frames.

It says in the \ConTeXt\ reference manual that \TeX\ has a huge text
management capability, but is very weak in managing graphic information.  I
beg to differ: it is true that for handling lines and frames the
possibilities of \ConTeXt\ (actually \TeX) are not as overwhelming as when
it comes to typesetting text.  But to go on to say that the system is weak
in this regard is, I think, somewhat of a stretch.  I don't know of any
function with lines and frames that other typesetting systems can do for
documents that \ConTeXt\ is unable to generate.  And if we
combine \ConTeXt\ with MetaPost, or with TiKZ (\ConTeXt\ has an expansion
module for this), then the possibilities are only limited by our
imagination.

In the following sections, however, I will limit myself to explaining how
to generate simple horizontal and vertical lines and frames around words,
sentences or paragraphs.

\startsubsection
    [title={Simple lines}]

The simplest way of drawing a horizontal line is with the
\PlaceMacro{hairline}\tex{hairline} command that generates a horizontal
line that occupies all the width of a normal text line.

Because \tex{hairline} occupies the entire width of a text line, there is
no room for any text on the same line as the \tex{hairline}. In order to
generate a horizontal line capable of coexisting with text, we need
the \PlaceMacro{thinrule}\tex{thinrule} command.  This second command will
use the full width of the line if it is used between paragraphs; in such a
case it will have the same effect as \tex{hairline}.  When used in a
paragraph, \tex{thinrule} will produce the same horizontal expansion
as \tex{hfill} (see \in{Section}[sec:horizontal space2]), but instead of
filling the horizontal space with white space (as \tex{hfill} does), it
fills it with a line.

\startbuffer
On the left\thinrule\\
\thinrule On the right\\
On both\thinrule sides\\
\thinrule centred\thinrule
\stopbuffer

\ShowAndExecute

With the \PlaceMacro{thinrules}\tex{thinrules} command we can generate
several lines.  For example, \tex{thinrules[n=2]} will generate two
consecutive lines, each the width of the normal line. The lines generated
with \tex{thinrules} can also be configured, either in an actual call to
the command, indicating the configuration as one of its arguments, or
generally with \tex{setupthinrules}. Configuration includes the thickness
of the line ({\tt rulethickness}), its colour ({\tt color}), background
colour ({\tt background}), interline space ({\tt interlinespace}), etc.

\startSmallPrint

    I will leave a number of options without explanation.  The reader can
    consult them in {\tt setup-en.pdf} (see
    \in{Section}[sec:qrc-setup-en]).  Some options only differ from others
    in terms of nuance (i.e., there is hardly any difference between them),
    and I think it is faster for the reader to try to {\em see} the
    difference, than for me to try to convey it in words.  For example: the
    thickness of the line I just said depends on the {\tt rulethickness}
    option.  But it is also affected by the {\tt height} and {\tt depth}
    options.

\stopSmallPrint

Smaller lines can be generated with the \PlaceMacro{hl}\tex{hl} and
\PlaceMacro{vl}\tex{vl} commands.  The first generates a horizontal line
and the second a vertical line.  Both take a number as a parameter that
allows us to specify the length of the line.  In \tex{hl} the number
measures the length in {\em ems} (no need to indicate the unit of
measurement in this command) and in \tex{vl} the argument is multiplied by
the current height of the text line to obtain the line length.

Thus \tex{hl[3]} generates a horizontal line of 3 ems and \tex{vl[3]}
generates a vertical line of the height corresponding to three
lines.  Remember that the line measurement indicator must be inserted
between square brackets, not between curly brackets.  In both commands the
argument is optional.  If it is not entered, a value of 1 is assumed.  Here
is \tex{vl[1]}: \vl[1].

\PlaceMacro{fillinline}\tex{fillinline} is another command to create
horizontal lines of precise length. It supports more configuration in which
we can indicate (or predetermine with
\PlaceMacro{setupfillinlines}\tex{setupfillinlines}) the width ({\tt width}
option) in addition to some other features.

A peculiarity of this command is that text that is written to its right
will be placed on the left of the line, separating that text from the line
by the necessary white space to occupy the whole line. For example:

\starttyping
\fillinline[width=6cm] Name
\stoptyping

will generate

\startcolor[red]

\fillinline[width=6cm] {Name}

\stopcolor

\startSmallPrint

    I suspect that this strange operation is due to the fact that this
    macro was designed to write forms where there is a horizontal line
    following the text, on which something \Conjecture must be written.
    % TODO: do we want this in the English version?
    % (The explanation makes sense for the Spanish version.)
    % In fact the very name of the command {\tt fillinline}
    % means \quotation{fill in the line}.

\stopSmallPrint

In addition to the width of the line, we can configure the margin ({\tt
    margin}), the distance ({\tt distance}), the thickness ({\tt
    rulethickness}) and the colour ({\tt color}).

Almost identical to \tex{fillinline} is \tex{fillinrules}, although this
command allows us to insert more than one line (\MyKey{n} option):

\type{\fillinrules [Configuration] {Text before rule} {Text after rule}}

where the three arguments are optional.  (To get text after the rule, use
two sets of curly brackets, where the first set is empty.)  For example, the line

\type{\fillinrules[distance=1cm, separator=?, n=2]{To be} {or not to be}}

\startcolor[red]
\fillinrules[distance=0.5cm, separator=?, n=2]{To be} {or not to be}
\stopcolor

Warning: putting a space between the close square bracket and the first
open curly bracket results in something considerably different:

\startcolor[red]
\fillinrules[distance=0.5cm, separator=?, n=2] {To be} {or not to be}
\stopcolor

\stopsubsection


\startsubsection
  [title={Lines linked to text}]

Although some of the commands we have just seen can generate lines that
coexist with text on the same line, those commands actually focus on
drawing the line.  To write lines linked to certain text, \ConTeXt\ has
commands that:

\startitemize

\item Generate text between lines.

\item Generate lines under the text (underlining), above the text
(overlining) or through the text (strikethrough).

\stopitemize

To generate a text between lines the usual command is
\PlaceMacro{textrule}\tex{textrule}.  This command draws a line that crosses
the entire width of the page and writes the text it takes as a parameter on
the left side (but not at the margin).  For example:

\startbuffer
\textrule{Example text}
\stopbuffer
\ShowAndExecute

\startSmallPrint

    It is assumed that \tex{textrule} allows an optional first argument
    with three possible values: {\tt top}, {\tt middle} and \Doubt{\tt
    bottom}.  But, after some tests, I have not been able to find out
    what effect such options produce.

\stopSmallPrint

Similar to \tex{textrule} is the
\PlaceMacro{starttextrule}\tex{starttextrule} environment which, besides
inserting the line with text at the beginning of the environment, inserts a
horizontal line at the end. The format of this command is:

\type{\starttextrule[Configuration]{Text on the line} ... \stoptextrule}

% TODO: why is there a blank line (or big vertical space) before the lower hrule here?

\startbuffer
\starttextrule{Intro text}
Some text which will get set
between two lines.  The bottom
line goes across the page.
\stoptextrule
\stopbuffer

\ShowAndExecute

Both \tex{textrule} and \text{starttextrule} can be configured with
\PlaceMacro{setuptextrule}\tex{setuptextrule}.

To draw lines under, over, or through text, the following commands are used:

\PlaceMacro{underbar}\PlaceMacro{underbars}\PlaceMacro{overbar}
\PlaceMacro{overbars}\PlaceMacro{overstrike}\PlaceMacro{overstrikes}

\startbuffer[buf1]
  \underbar{Text words}   
  \underbars{Text words}
  \overbar{Text words}
  \overbars{Text words}
  \overstrike{Text words}
  \overstrikes{Text words}
\stopbuffer
\startbuffer[buf2]
\startlines
  \getbuffer[buf1]
\stoplines
\stopbuffer

% In this example increase the line spacing to make the rules easier to see.
% But counteract the extra space between the line above and this example.
% TODO: what is The Right ConTeXt Way to do this?
% (Part of the issue is that depending on the rest of the page, the gap
% here changes size... the -0.44ex was correct when I put a \page right
% after the \stop here.)
\start
\godown[-1.5ex]\setupinterlinespace[line=3.55ex]
\ShowAndExecute[buf1][buf2]
\stop

As we can see, for each type of line (under, over, or through text) there
are two commands. The singular version of the command draws a single line
under, over or through all the text taken as an argument, while the plural
version of the command only draws the line under, over or through the
words, but not the white space.

These commands are not compatible with each other; that is to say, two of
them cannot be applied to the same text.  If we try, the last one will
always prevail.  On the other hand, \tex{underbar} can be repeated, underlining
what has already been underlined: \type{\underbar{\underbar{two underbars}}} 
outputs \underbar{\underbar{two underbars}}.

% TODO: test the following bit out and rewrite.

\startSmallPrint

    The reference manual points out that \tex{underbar} disables
    hyphenation of words in the text that constitute its argument.  It is
    not clear to me whether that statement refers only to \tex{underbar} or
    to the six commands we are examining.

\stopSmallPrint

\stopsubsection


\startsubsection
  [title={Framed words or texts}]

To surround a text with a frame (i.e., a box) we use:

\startitemize

  \item The \PlaceMacro{framed}\tex{framed} or
    \PlaceMacro{inframed}\tex{inframed} commands if the text is relatively
    brief and does not take up more than one line.

  \item The \PlaceMacro{startframedtext}\tex{startframedtext} environment
    for longer texts.

\stopitemize

The difference between \tex{framed} and \tex{inframed} lies in the point
from which the frame is drawn.  In \tex{frame} the frame is drawn upwards
from an ideal line, called a baseline, on which the letters rest, but
certain letters pass downwards (letters like \quote{j} and \quote{p} have
{\em descenders}, i.e., strokes that extend below the baseline).
In \tex{inframed} the frame is drawn, also upwards, from the lowest
possible point on the line.  For example:

\startbuffer
Here there are \framed{two} good
\inframed{frames}.
\stopbuffer
\ShowAndExecute

Both \tex{framed} and \tex{inframed} text can be customized with
\PlaceMacro{setupframed}\tex{setupframed}, and \tex{startframedtext} is
customized with \PlaceMacro{setupframedtext}\tex{setupframedtext}. The
customization options for both commands are quite similar.  They allow us
to indicate the measurements of the frame ({\tt height}, {\tt width}, {\tt depth}), the
shape of the corners ({\tt framecorner}), which can be {\tt rectangular} or
round ({\tt round}), the frame colour ({\tt framecolor}), the line
thickness ({\tt framethickness}), the alignment of content ({\tt align}),
text colour ({\tt foregroundcolor}), background colour ({\tt background}
and {\tt backgroundcolor}), etc.

% TODO: I think an example or two here would be useful.

For \tex{startframedtext} there is also an apparently strange option, {\tt
    frame=off}, that causes the frame not to be drawn (although it is still
there, but invisible).  This option is useful in cases where a page break
within a paragraph (or other material which should stay together) would be
undesired.  Since the frame around a paragraph is indivisible, enclosing a
paragraph in a {\tt framedtext} environment (even with the frame drawing
option turned off) ensures that no page breaks are inserted within the
paragraph.

We can also create a customized version of these commands with
\PlaceMacro{defineframed}\tex{defineframed} and
\PlaceMacro{defineframedtext}\tex{defineframedtext}.

\stopsubsection

\stopsection


\startsection
  [
    reference=sec:buffer,
    title={Other environments and constructions of interest},
  ]

  There are still many environments in \ConTeXt\ that I have not even
  mentioned, or have only done so in passing.  By way of example:

\startitemize

  \item {\tt\bf buffer}\PlaceMacro{startbuffer}\PlaceMacro{getbuffer}: {\em
    Buffers} are text fragments stored in memory for later re-use. A {\em
    buffer} is defined somewhere in the document with
    \cmd{startbuffer[BufferName] ... \ttbs stopbuffer} and can be
    retrieved as often as desired at some other point in the document with
    \tex{getbuffer[BufferName]}.  If you examine the \quotation{source
        code} of this chapter, you will see that buffers are frequently
    used in the side-by-side examples, so that the text which is displayed
    (with \tex{typebuffer[BufferName]}) and typeset
    (with \tex{getbuffer[BufferName]}) only has to be entered once.

  \item {\tt\bf chemical}\PlaceMacro{startchemical}: This environment allows
    us to place chemical formulas inside it. If \TeX\ stands out, among
    many other things, for its ability to typeset texts with mathematical
    formulas properly, from the outset \ConTeXt\ sought to extend this
    ability to chemical formulas, and it has this environment where
    commands and structures are enabled for writing chemical formulas.

  \item {\tt\bf combination}\PlaceMacro{startcombination}: This environment
    allows us to combine several floating elements
    (see \in{Section}[sec:floating objects]) on the same page.  It is
    particularly useful for aligning different connected external images in
    our document.

  \item {\tt\bf formula}\PlaceMacro{startformula}: This is an environment
    aimed at typesetting maths formulas.

  \item {\tt\bf hiding}\PlaceMacro{starthiding}: The text stored in this
    environment will not be compiled and therefore will not appear in the
    final document.  This is useful for temporarily disabling compilation of
    certain fragments in the source file.  The same thing is achieved by
    marking one or more lines as a comment.  But when the fragment we want
    to disable is relatively long, more effective than marking tens or
    hundreds of lines of the source file as a comment is to insert
    the \tex{starthiding} command at the beginning of the fragment,
    and \tex{stophiding} at the end.

  \item {\tt\bf legend}\PlaceMacro{startlegend}: In a mathematical
    context, \TeX\ applies different rules so that no normal text can be
    written.  However, sometimes a formula is accompanied by a description
    of the elements used in it.  For this purpose there is
    the \tex{startlegend} environment which allows us to place normal text
    in a mathematical context.

  \item {\tt\bf linecorrection}\PlaceMacro{startlinecorrection}:
    Usually, \ConTeXt\ correctly manages the vertical space between lines,
    but occasionally a line may contain something that causes it to not
    display correctly.  This happens mainly with lines that have fragments framed
    with \tex{framed}.  In such cases this environment adjusts the line spacing
    so that the paragraph appears correctly.

  \item {\tt\bf mode}\PlaceMacro{startmode}: This environment is intended
    to include fragments in the source file that will only be compiled if
    the appropriate mode is active.  The use of {\em modes} is not the
    subject of this introduction, but it is a very interesting tool if we
    want to be able to generate several versions of our document with
    different formats (e.g., a version to be printed or a version optimized
    for reading on a screen), all from a single source file.  A complementary
    environment to this one is \PlaceMacro{startnotmode}\tex{startnotmode},
    where source material enclosed between \tex{startnotmode[someMode]}
    and \tex{stopnotmode} is included in the document when the mode {\tt
    someMode} has not been activated.

    % TODO: There is a nice little example in
    % fr2/02-03_BodyMatter_MainFlow/20_01_Markup_Others.tex 
    % lines 229-266.  Include that (with permission) maybe something like it.

  \item {\tt\bf opposite}\PlaceMacro{startopposite}: This environment is
    used to typeset texts when the contents of the left and right pages are
    related.

  \item {\tt\bf quotation}\PlaceMacro{startquotation}: A very similar
    environment to {\tt narrower} (see \in{Section}[sec:narrower]),
    intended to insert moderately long literal quotations.  This
    environment makes sure that the text inside is quoted, and that the
    margins are increased so that the paragraph with the quotation stands
    out visually on the page.  But it should be noted that according to
    common block-quote styles in English, there should be no opening and
    closing quotation marks --- which makes this command or environment
    less useful than it might be otherwise.

  \item {\tt\bf standardmakeup}\PlaceMacro{startstandardmakeup}: This
    environment is designed to generate pages with the title of the
    document, which is relatively common in academic documents of a certain
    length, such as doctoral theses, master's theses, etc.

\stopitemize

To learn about any of these environments (or others I have not mentioned),
I suggest the following steps:

\startitemize[n]

  \item Look for information on the environment in the \ConTeXt\ reference
    manual. This manual does not mention all the environments; but it does
    say something about every item in the list above.

  \item Write a test document where the environment is used.

  \item Look up \ConTeXt's official list of commands (see
    \in{Section}[sec:qrc-setup-en]) for the configuration options for the
    environment in question, then test them to see exactly what they do.

\stopitemize

\stopchapter

\stopcomponent

%%% Local Variables:
%%% mode: ConTeXt
%%% eval: (auto-fill-mode 1)
%%% coding: utf-8-unix
%%% TeX-master: "../introCTX_eng.tex"
%%% End:
%%% vim:set filetype=context tw=75 : %%%
